---
title: "Descriptive Analysis"
output:
  html_document:
    toc: true
    toc_depth: 3
    number_sections: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
knitr::opts_chunk$set(fig.path = "figures/", fig.width = 8, fig.height = 5)
```

```{r fixlib}

# Rcpp is too new after renewing RStudio

remove.packages("Rcpp")
install.packages("Rcpp", version = "1.0.12")

```

# 1. Preparation

```{r libraries}
options(repos = c(CRAN = "https://mirrors.tuna.tsinghua.edu.cn/CRAN/"))
# packages
install.packages("dplyr")
install.packages("stringr")
install.packages("ggplot2")
install.packages("tidyr")
install.packages("readxl")
install.packages("scales")
install.packages("showtext")
install.packages("sysfonts")
install.packages("skimr")
install.packages("plm")
install.packages("car")
install.packages("glmnet")
install.packages("Rtsne")
install.packages("networkD3")
install.packages("cluster")
install.packages("factoextra")
install.packages("pROC")
install.packages("lme4")
install.packages("MuMIn")
install.packages("nnet")
install.packages("DHARMa")
install.packages("tigris")
install.packages("ggeffetcs")
install.packages("stargazer")
install.packages("htmltools")
install.packages("tidyverse")
install.packages("ggcorrplot")
install.packages("purr")
install.packages("broom.mixed")
install.packages("performance") 
install.packages("lattice")
install.packages("fmsb")
install.packages("kableExtra")
install.packages("broom")
install.packages("knitr")
install.packages("ggnewscale") 
install.packages("modelsummary")
install.packages("pscl")
install.packages("ggfortify")
install.packages("rlang")
install.packages("patchwork")
install.packages("segmented")
install.packages("tidytext")
install.packages("mgcv")
install.packages("splines")
install.packages("interactions")


library(tidytext)
library(segmented)
library(pscl)
library(dplyr)
library(stringr)
library(ggplot2)
library(tidyr)
library(readxl)
library(scales) 
library(showtext)
library(stringr)
library(sysfonts)
library(skimr)
library(plm)
library(car)
library(Rtsne)
library(networkD3)
library(glmnet)
library(fmsb)
library(cluster)
library(factoextra)
library(pROC)
library(lme4)
library(MuMIn)
library(nnet)
library(DHARMa)
library(tigris)
library(sf)
library(cowplot)
library(ggeffects)
library(stargazer)
library(htmltools)
library(tidyverse)
library(ggcorrplot)
library(purrr)
library(broom.mixed)
library(performance) 
library(lattice)
library(fmsb)
library(kableExtra)
library(broom)
library(knitr)
library(ggnewscale)
library(modelsummary)
library(ggfortify)
library(rlang)
library(patchwork)
library(mgcv)
library(splines)
library(interactions)
```

# 2. basic data cleansing

```{r load-data}

# ------ data acquiry ------
adver_sum_path <- "/Users/kunyazhan/Desktop/data/0901data/adver_analysis_sum.xlsx"
legis_sum_path <- "/Users/kunyazhan/Desktop/data/0901data/legis_analysis_sum.xlsx"
adver_sum <- read_excel(adver_sum_path)
legis_sum <- read_excel(legis_sum_path)

adver_path <- "/Users/kunyazhan/Desktop/data/0901data/adver_analysis_all.xlsx"
legis_path <- "/Users/kunyazhan/Desktop/data/0901data/legis_analysis_all.xlsx"
adver_all <- read_excel(adver_path)
legis_all <- read_excel(legis_path)



adver_data <- read_csv("/Users/kunyazhan/Desktop/data/0901data/adver_train_data.csv")
legis_data <- read_csv("/Users/kunyazhan/Desktop/data/0901data/legis_train_data.csv")

match_vars <- c("leverage_pl", "leverage_ne", "committee_pl", "committee_ne",
                "employ_industry_pl", "employ_industry_ne",
                "payroll_industry_pl", "payroll_industry_ne",
                "ethnicity_pl", "ethnicity_ne",
                "party_value_pl", "party_value_ne",
                "party_score_pl", "party_score_ne")


adver_data_raw <- adver_data %>%
  rename_with(.cols = all_of(match_vars), .fn = ~ paste0(.x, "_raw"))

adver_all <- adver_all %>%
  left_join(adver_data_raw %>% select(full_name, issue, congress, ends_with("_raw")),
            by = c("full_name", "issue", "congress"))

legis_data_raw <- legis_data %>%
  rename_with(.cols = all_of(match_vars), .fn = ~ paste0(.x, "_raw"))

legis_all <- legis_all %>%
  left_join(legis_data_raw %>% select(full_name, issue, congress, ends_with("_raw")),
            by = c("full_name", "issue", "congress"))

# match features
stopifnot(all(paste0(match_vars, "_raw") %in% names(adver_all)))
stopifnot(all(paste0(match_vars, "_raw") %in% names(legis_all)))
```

```{r quick-glance}
# a quick glance for whole picture
skim(adver_sum)
skim(legis_sum)


#specifically, a quick glance at intl/dom/prof importance
summary(legis_sum$importance_intl)
summary(legis_sum$importance_dom)
summary(legis_sum$importance_prof)
summary(legis_sum$importance_all)

summary(adver_sum$importance_intl)
summary(adver_sum$importance_dom)
summary(adver_sum$importance_prof)
summary(adver_sum$importance_all)
```

# 3. overall description for decision model - each legislators

```{r clean-and-replace-ideology}


# data cleansing 

bioguide_path <- "/Users/kunyazhan/Desktop/data/new ideology/bioguide.xlsx"
bioguide <- read_excel(bioguide_path)

chamber.ideology <- read_csv("/Users/kunyazhan/Desktop/data/new ideology/chamber ideology.csv")


# 第一步：将 bioguide_id 从 bioguide 合并进 adver_sum（通过 full_name + congress） ----
adver_sum <- adver_sum %>%
  left_join(
    bioguide %>% select(full_name, congress, bioguide_id),
    by = c("full_name", "congress")
  )
legis_sum <- legis_sum %>%
  left_join(
    bioguide %>% select(full_name, congress, bioguide_id),
    by = c("full_name", "congress")
  )

adver_missing <- adver_sum %>% 
  filter(is.na(bioguide_id)) # missing
legis_missing <- legis_sum %>% 
  filter(is.na(bioguide_id)) # no missing

# 解决missing问题

bioguide_name_map <- bioguide %>%
  select(full_name, bioguide_id) %>%
  distinct(full_name, .keep_all = TRUE)

adver_missing_fixed <- adver_missing %>%
  left_join(bioguide_name_map, by = "full_name", suffix = c("", ".new")) %>%
  mutate(
    bioguide_id = ifelse(is.na(bioguide_id), bioguide_id.new, bioguide_id)
  ) %>%
  select(-bioguide_id.new)


adver_sum_fixed <- adver_sum %>%
  left_join(adver_missing_fixed %>%
              select(full_name, congress, bioguide_id) %>%
              distinct(),
            by = c("full_name", "congress"),
            suffix = c("", "_new")) %>%
  mutate(bioguide_id = coalesce(bioguide_id, bioguide_id_new)) %>%
  select(-bioguide_id_new)


adver_missing <- adver_sum_fixed %>% 
  filter(is.na(bioguide_id))

adver_sum <- adver_sum %>%
  left_join(adver_missing_fixed %>%
              select(full_name, congress, bioguide_id) %>%
              distinct(),
            by = c("full_name", "congress"),
            suffix = c("", "_new")) %>%
  mutate(bioguide_id = coalesce(bioguide_id, bioguide_id_new)) %>%
  select(-bioguide_id_new)

adver_sum_old <- adver_sum
adver_sum <- adver_sum_fixed


# 第二步：用 bioguide_id + congress + chamber 合并 ideology 数据 ----

# 处理chamber.ideology

chamber.ideology <- chamber.ideology %>%
  mutate(
    congress = as.character(congress),
    chamber = as.character(chamber)
  )

chamber.ideology <- chamber.ideology %>%
  mutate(chamber = case_when(
    chamber == "house" ~ "House of Representatives",
    chamber == "senate" ~ "Senate",
    TRUE ~ chamber
  ))





# 匹配

adver_sum <- adver_sum %>%
  mutate(
    congress = as.character(congress),
    chamber = as.character(chamber)
  )
adver_sum <- adver_sum %>%
  left_join(
    chamber.ideology %>%
      select(bioguide_id, congress, chamber, ideology) %>%
      rename(ideology_new = ideology),
    by = c("bioguide_id", "congress", "chamber")
  )


legis_sum <- legis_sum %>%
  mutate(
    congress = as.character(congress),
    chamber = as.character(chamber)
  )
legis_sum <- legis_sum %>%
  left_join(
    chamber.ideology %>%
      select(bioguide_id, congress, chamber, ideology) %>%
      rename(ideology_new = ideology),
    by = c("bioguide_id", "congress", "chamber")
  )


adver_sum <- adver_sum %>%
  mutate(ideology = ideology_new) %>%
  select(-ideology_new)

legis_sum <- legis_sum %>%
  mutate(ideology = ideology_new) %>%
  select(-ideology_new)

# 查看缺失

adver_missing <- adver_sum %>% 
  filter(is.na(ideology))
legis_missing <- legis_sum %>% 
  filter(is.na(ideology))

unique(adver_missing$bioguide_id)
unique(legis_missing$bioguide_id)

```


```{r aggregate-data}


# data aggregation
generate_decision_with_traits <- function(df, mode = "all") {
  # 判断合法输入
  if (!mode %in% c("all", "train", "test")) {
    stop("mode 必须是 'all', 'train', 或 'test'")
  }

  # 根据 mode 筛选数据
  df_subset <- if (mode == "all") {
    df
  } else {
    df %>% filter(dataset == mode)
  }

  # 汇总每位议员在特定届别和数据集中的重要性得分
  df_clean <- df_subset %>%
    group_by(full_name, congress, dataset) %>%
    summarise(
      importance_intl = sum(importance_intl, na.rm = TRUE),
      importance_dom  = sum(importance_dom,  na.rm = TRUE),
      importance_prof = sum(importance_prof, na.rm = TRUE),
      importance_all  = sum(importance_all,  na.rm = TRUE),
      pred = mean(pred, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      intl_percentage = ifelse(importance_all > 0, importance_intl / importance_all, NA_real_),
      dom_percentage  = ifelse(importance_all > 0, importance_dom  / importance_all, NA_real_),
      prof_percentage = ifelse(importance_all > 0, importance_prof / importance_all, NA_real_),
      max_val = pmax(intl_percentage, dom_percentage, prof_percentage, na.rm = TRUE),
      decision = case_when(
        is.na(max_val) ~ NA_character_,
        intl_percentage == max_val & dom_percentage != max_val & prof_percentage != max_val ~ "intl",
        dom_percentage  == max_val & intl_percentage != max_val & prof_percentage != max_val ~ "dom",
        prof_percentage == max_val & intl_percentage != max_val & dom_percentage != max_val ~ "prof",
        TRUE ~ "tie"
      )
    ) %>%
    select(-max_val, -importance_all)

  # 合并特征信息
  traits <- df %>%
    select(full_name, congress, dataset, party, chamber, committee_el,
           gender, age, seniority, ideology, district, state_code, bioguide_id) %>%
    distinct()

  df_final <- df_clean %>%
    left_join(traits, by = c("full_name", "congress", "dataset")) %>%
    mutate(gender = factor(gender, levels = c(0, 1), labels = c("Female", "Male")))

  return(df_final)
}

# adv
adver_cg <- generate_decision_with_traits(adver_sum)

#df_train <- generate_decision_with_traits_dual(df, mode = "train")
#df_test <- generate_decision_with_traits_dual(df, mode = "test")

# legis
legis_cg <- generate_decision_with_traits(legis_sum)

# adver_cg + district features

# 转字符型避免 join 报错
adver_all_clean <- adver_all %>%
  mutate(congress = as.character(congress))

adver_cg_clean <- adver_cg %>%
  mutate(congress = as.character(congress))

# 只保留唯一行
adver_all_subset <- adver_all_clean %>%
  select(full_name, congress,
         employ1, employ2, employ3,
         payroll1, payroll2, payroll3,
         ethnicity1, ethnicity2, ethnicity3,
         ppvi, npvi) %>%
  distinct()  # 去重

# 合并
adver_cg_clean <- adver_cg_clean %>%
  left_join(adver_all_subset, by = c("full_name", "congress"))


# legis_cg + district features

# 转字符型避免 join 报错
legis_all_clean <- legis_all %>%
  mutate(congress = as.character(congress))

legis_cg_clean <- legis_cg %>%
  mutate(congress = as.character(congress))

# 只保留唯一行
legis_all_subset <- legis_all_clean %>%
  select(full_name, congress,
         employ1, employ2, employ3,
         payroll1, payroll2, payroll3,
         ethnicity1, ethnicity2, ethnicity3,
         ppvi, npvi) %>%
  distinct()  # 去重

# 合并
legis_cg_clean <- legis_cg_clean %>%
  left_join(legis_all_subset, by = c("full_name", "congress"))


```

## 3.1 distribution bar plot

```{r distribution_plot}

# adver
ggplot(adver_cg, aes(x = decision, fill = decision)) +
  geom_bar() +
  scale_fill_manual(
    values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c")
  ) +
  geom_text(stat='count', aes(label=..count..), vjust=-0.8, size=5) +
  labs(title = "decision mode by congressmen adver",
       x = "decision mode",
       y = "number") +
  theme_minimal(base_size = 14)


#legis
ggplot(legis_cg, aes(x = decision, fill = decision)) +
  geom_bar() +
  scale_fill_manual(
    values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c")
  ) +
  geom_text(stat='count', aes(label=..count..), vjust=-0.8, size=5) +
  labs(title = "decision mode by congressmen legis",
       x = "decision mode",
       y = "number") +
  theme_minimal(base_size = 14)


# barplot by each congress
ggplot(adver_cg, aes(x = decision, fill = decision)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.0, size = 3) +
  scale_fill_manual(
    values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c")
  ) +
  facet_wrap(~ congress, scales = "free_y") +
  labs(title = "decision mode by congressmen adver",
       x = "Decision Mode",
       y = "Number") +
  theme_minimal(base_size = 12)

ggplot(legis_cg, aes(x = decision, fill = decision)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = after_stat(count)), vjust = -0.0, size = 3) +
  scale_fill_manual(
    values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c")
  ) +
  facet_wrap(~ congress, scales = "free_y") +
  labs(title = "decision mode by congressmen legis",
       x = "Decision Mode",
       y = "Number") +
  theme_minimal(base_size = 12)


# label's distribution 

table(adver_all$congress[adver_all$label == 0])
table(legis_all$congress[legis_all$label == 0])
table(adver_all$party[adver_all$label == 1])
table(legis_all$party[legis_all$label == 1])
table(adver_all$chamber[adver_all$label == 1])
table(legis_all$chamber[legis_all$label == 1])



```

## 3.2 change of decision mode

```{r decision-mode}

# how many legislators changed their decision mode
changed_members <- adver_cg %>%
  group_by(full_name) %>%
  summarise(n_decisions = n_distinct(decision)) %>%
  filter(n_decisions > 1)

changed_members1 <- legis_cg %>%
  group_by(full_name) %>%
  summarise(n_decisions = n_distinct(decision)) %>%
  filter(n_decisions > 1)

print(changed_members, n = Inf) 
print(changed_members1, n = Inf) 


# Aggregate both adver_cg and legis_cg to see how they are between 2 behavior

adver_mode_df <- adver_cg %>%
  select(full_name, congress, adver_mode = decision)

legis_mode_df <- legis_cg %>%
  select(full_name, congress, legis_mode = decision)

merged_df <- inner_join(adver_mode_df, legis_mode_df, by = c("full_name", "congress"))

sankey_data <- merged_df %>%
  count(adver_mode, legis_mode, name = "value") %>%
  mutate(
    adver_mode = paste0("adver-", adver_mode),
    legis_mode = paste0("legis-", legis_mode)
  )

# nodes
nodes <- data.frame(name = unique(c(sankey_data$adver_mode, sankey_data$legis_mode)))

nodes <- data.frame(name = c(
  "adver-intl", "adver-dom", "adver-prof",
  "legis-intl", "legis-dom", "legis-prof"
))
nodes$group <- gsub(".*-", "", nodes$name)

sankey_data <- sankey_data %>%
  mutate(
    source = match(adver_mode, nodes$name) - 1,
    target = match(legis_mode, nodes$name) - 1
  )

# constructing color scale
my_color <- 'd3.scaleOrdinal()
  .domain(["intl", "dom", "prof"])
  .range(["#1f77b4", "#ff7f0e", "#2ca02c"])'

# sankey plot
sankey_plot <- sankeyNetwork(Links = sankey_data,
              Nodes = nodes,
              Source = "source",
              Target = "target",
              Value = "value",
              NodeID = "name",
              NodeGroup = "group",     
              colourScale = my_color,  
              fontSize = 14,
              nodeWidth = 30,
              sinksRight = FALSE)

saveNetwork(sankey_plot, "sankey_plot.html", selfcontained = TRUE)

```

## 3.3 committee decision relationship

```{r decision-by-committee}

# under legis 

# unique committee
legis_cg %>% 
  distinct(committee_el) %>% 
  nrow()

unique(legis_cg$committee_el)


# ======= decision mode percentage for legislators each committee ======

committee_decision_summary <- legis_cg %>%
  filter(!is.na(decision)) %>%
  group_by(committee_el, decision) %>%
  summarise(count = n(), .groups = "drop") %>%
  group_by(committee_el) %>%
  mutate(
    total = sum(count),
    percentage = count / total
  ) %>%
  arrange(committee_el, desc(percentage))

print(committee_decision_summary)

avg_mode_percentage_all <- committee_decision_summary %>%
  group_by(decision) %>%
  summarise(avg_percentage = mean(percentage, na.rm = TRUE)) %>%
  arrange(desc(avg_percentage))


print(avg_mode_percentage_all)


target_committees <- c(
  "Committee on Armed Services",
  "Committee on Foreign Affairs",
  "Committee on Foreign Relations",
  "Committee on Homeland Security",
  "Committee on Homeland Security and Governmental Affairs",
  "Committee on Science, Space, and Technology",
  "Permanent Select Committee on Intelligence"
)

target_mode_distribution <- committee_decision_summary %>%
  filter(committee_el %in% target_committees) %>%
  select(committee_el, decision, percentage) %>%
  arrange(committee_el, desc(percentage))

print(target_mode_distribution)


# ------- intl + prof as a whole mode percentage --------

# overall summary for all committees
overall_summary <- committee_decision_summary %>%
  group_by(decision) %>%
  summarise(total_count = sum(count), .groups = "drop") %>%
  mutate(
    total_all = sum(total_count),
    overall_percentage = total_count / total_all
  )

#  dom per and intl+prof per for overall committees
dom_ratio <- overall_summary %>% filter(decision == "dom") %>% pull(overall_percentage)
intl_prof_ratio <- overall_summary %>% 
  filter(decision %in% c("intl", "prof")) %>% 
  summarise(sum_ratio = sum(overall_percentage)) %>% 
  pull(sum_ratio)

# output results
cat(" Overall average proportions across all committees:\n")
cat(sprintf("→ dom: %.2f%%\n", dom_ratio * 100))
cat(sprintf("→ intl + prof: %.2f%%\n", intl_prof_ratio * 100))

# targeted committees
target_committees <- c(
  "Committee on Armed Services",
  "Committee on Foreign Affairs",
  "Committee on Foreign Relations",
  "Committee on Homeland Security",
  "Committee on Homeland Security and Governmental Affairs",
  "Committee on Science, Space, and Technology",
  "Permanent Select Committee on Intelligence"
)

# decision mode perentage for targeted committees
target_summary <- committee_decision_summary %>%
  filter(committee_el %in% target_committees) %>%
  select(committee_el, decision, percentage) %>%
  arrange(committee_el, desc(percentage))

print(target_summary)

# dom vs intl+prof 
target_summary_collapsed <- committee_decision_summary %>%
  filter(committee_el %in% target_committees) %>%
  mutate(
    decision_group = ifelse(decision == "dom", "dom", "intl_prof")
  ) %>%
  group_by(committee_el, decision_group) %>%
  summarise(group_percentage = sum(percentage), .groups = "drop") %>%
  pivot_wider(
    names_from = decision_group,
    values_from = group_percentage,
    values_fill = 0
  ) %>%
  arrange(desc(intl_prof))  # ranks as wanted

print(target_summary_collapsed)



# dom vs intl+prof for all committess

committee_summary_collapsed <- committee_decision_summary %>%
  mutate(
    decision_group = ifelse(decision == "dom", "dom", "intl_prof")
  ) %>%
  group_by(committee_el, decision_group) %>%
  summarise(group_percentage = sum(percentage), .groups = "drop") %>%
  pivot_wider(
    names_from = decision_group,
    values_from = group_percentage,
    values_fill = list(group_percentage = 0)
  ) %>%
  
  mutate(
    intl_prof_raw = intl_prof,
    dom = sprintf("%.2f%%", dom * 100),
    intl_prof = sprintf("%.2f%%", intl_prof * 100)
  ) %>%
  arrange(desc(intl_prof_raw)) %>%
  select(committee_el, dom, intl_prof)


print(committee_summary_collapsed, n = Inf)

```

## 3.4 issue rank by each legislator
```{r issue-rank-by-pred}

# data cleansing

generate_decision_with_issue <- function(df, set_name = "train") {
  # 1. 按 full_name 和 congress 分组，对 importance 求和，pred 求平均
  df_clean <- df %>%
    filter(dataset == set_name) %>%
    group_by(full_name, congress, issue) %>%
    summarise(
      importance_intl = sum(importance_intl, na.rm = TRUE),
      importance_dom = sum(importance_dom, na.rm = TRUE),
      importance_prof = sum(importance_prof, na.rm = TRUE),
      pred = mean(pred, na.rm = TRUE),
      label = sum(label, na.rm = TRUE),
      .groups = "drop"
    ) %>%
    mutate(
      total = importance_intl + importance_dom + importance_prof,
      intl_percentage = ifelse(total > 0, importance_intl / total, NA_real_),
      dom_percentage = ifelse(total > 0, importance_dom / total, NA_real_),
      prof_percentage = ifelse(total > 0, importance_prof / total, NA_real_),
      max_val = pmax(intl_percentage, dom_percentage, prof_percentage, na.rm = TRUE),
      decision = case_when(
        is.na(max_val) ~ NA_character_,
        intl_percentage == max_val & dom_percentage != max_val & prof_percentage != max_val ~ "intl",
        dom_percentage == max_val & intl_percentage != max_val & prof_percentage != max_val ~ "dom",
        prof_percentage == max_val & intl_percentage != max_val & dom_percentage != max_val ~ "prof",
        TRUE ~ "tie"
      )
    ) %>%
    select(-max_val, -total)

  # 2. 提取个人特征变量，distinct去重
  traits <- df %>%
    filter(dataset == set_name) %>%
    select(full_name, congress, party, chamber, committee_el, gender, age, seniority, ideology) %>%
    distinct()

  # 3. left_join合并
  df_final <- df_clean %>%
    left_join(traits, by = c("full_name", "congress")) %>%
    mutate(gender = factor(gender, levels = c(0, 1), labels = c("Female", "Male")))

  return(df_final)
}

adver_cg_tp <- generate_decision_with_issue(adver_sum, set_name = "train")
legis_cg_tp <- generate_decision_with_issue(legis_sum, set_name = "train")

# ========= function: 1. get_top_issues_by_decision() =========

# This section defines a function is used for identify the most commonly prioritized policy issues (high pred) for legislators,
# categorized by their decision-making modes 

get_top_issues_by_decision <- function(df, top_n = 5) {
  # Step 1: for each legislator, select the issue with the highest predicted probability
  top_issue <- df %>%
    filter(!is.na(decision), decision %in% c("intl", "dom", "prof")) %>%
    group_by(full_name, congress) %>%
    filter(pred == max(pred, na.rm = TRUE)) %>%
    slice(1) %>%
    ungroup()

  # Step 2: count how frequently each issue as the top-predicted issue under each behavior mode
  issue_distribution <- top_issue %>%
    group_by(decision, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(decision) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  # Step 3: extract the top n issues for each decision mode
  top_issues <- issue_distribution %>%
    group_by(decision) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(decision, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_adver <- get_top_issues_by_decision(adver_cg_tp, top_n = 10)
result_legis <- get_top_issues_by_decision(legis_cg_tp, top_n = 10)

# print ranks ranks for all issues
print(result_adver$full_distribution)
print(result_legis$full_distribution)

# print top issues for each mode
print(result_adver$top_issues)
print(result_legis$top_issues)


# ========= function: 2. get_top_issues_by_decision_congress() ==========
# this section defines a function used to count the most frequently predicted issues chosen by legislators
# in each congress and decision mode

get_top_issues_by_decision_congress <- function(df, top_n = 5) {

  top_issue <- df %>%
    filter(!is.na(decision), decision %in% c("intl", "dom", "prof")) %>%
    group_by(full_name, congress) %>%
    filter(pred == max(pred, na.rm = TRUE)) %>%
    slice(1) %>%
    ungroup()

  
  issue_distribution <- top_issue %>%
    group_by(congress, decision, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(congress, decision) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  
  top_issues <- issue_distribution %>%
    group_by(congress, decision) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(congress, decision, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_adver_congress <- get_top_issues_by_decision_congress(adver_cg_tp, top_n = 3)
result_legis_congress <- get_top_issues_by_decision_congress(legis_cg_tp, top_n = 3)

# overall distribution by congress
print(result_adver_congress$full_distribution)
print(result_legis_congress$full_distribution)

# top issues of each mode by congress
print(result_adver_congress$top_issues)
print(result_legis_congress$top_issues)


# ========= function: 3. get_top_issues_by_party() ========
# this section defines a function used to count the most frequently predicted issues chosen by legislators
# in each party 

get_top_issues_by_party <- function(df, top_n = 5) {
  top_issue <- df %>%
    filter(!is.na(party)) %>%
    group_by(full_name, congress) %>%
    filter(pred == max(pred, na.rm = TRUE)) %>%
    slice(1) %>%
    ungroup()

  issue_distribution <- top_issue %>%
    group_by(party, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(party) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  top_issues <- issue_distribution %>%
    group_by(party) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(party, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_pred_party_adver <- get_top_issues_by_party(adver_cg_tp, top_n = 10)
result_pred_party_legis <- get_top_issues_by_party(legis_cg_tp, top_n = 10)

print(result_pred_party_adver$full_distribution)
print(result_pred_party_legis$full_distribution)

print(result_pred_party_adver$top_issues)
print(result_pred_party_legis$top_issues)


# ======== function: 4. get_top_issues_by_chamber() =======
# this section defines a function used to count the most frequently predicted issues chosen by legislators
# in each chamber


get_top_issues_by_chamber <- function(df, top_n = 5) {
 
  top_issue <- df %>%
    filter(!is.na(chamber)) %>%
    group_by(full_name, congress) %>%
    filter(pred == max(pred, na.rm = TRUE)) %>%
    slice(1) %>%
    ungroup()

  issue_distribution <- top_issue %>%
    group_by(chamber, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(chamber) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  top_issues <- issue_distribution %>%
    group_by(chamber) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(chamber, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_pred_chamber_adver <- get_top_issues_by_chamber(adver_cg_tp, top_n = 10)
result_pred_chamber_legis <- get_top_issues_by_chamber(legis_cg_tp, top_n = 10)

print(result_pred_chamber_adver$full_distribution)
print(result_pred_chamber_legis$full_distribution)

print(result_pred_chamber_adver$top_issues)
print(result_pred_chamber_legis$top_issues)



# ======= function: 5. get_top_issues_by_chamber_and_decision() =======
# this section defines a function used to count the most frequently predicted issues chosen by legislators
# in each chamber and decision mode

get_top_issues_by_chamber_and_decision <- function(df, top_n = 5) {

  top_issue <- df %>%
    filter(!is.na(chamber), !is.na(decision), decision %in% c("intl", "dom", "prof")) %>%
    group_by(full_name, congress) %>%
    filter(pred == max(pred, na.rm = TRUE)) %>%
    slice(1) %>%
    ungroup()

  issue_distribution <- top_issue %>%
    group_by(chamber, decision, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(chamber, decision) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  top_issues <- issue_distribution %>%
    group_by(chamber, decision) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(chamber, decision, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_pred_chamber_adver <- get_top_issues_by_chamber_and_decision(adver_cg_tp, top_n = 3)
result_pred_chamber_legis <- get_top_issues_by_chamber_and_decision(legis_cg_tp, top_n = 3)

print(result_pred_chamber_adver$full_distribution)
print(result_pred_chamber_legis$full_distribution)

print(result_pred_chamber_adver$top_issues)
print(result_pred_chamber_legis$top_issues)



```

```{r issue-rank-by-label}

# ======== function: 1. get_top_labeled_issues_by_decision() ========
# this section defines a function used to identify the most commonly emphasized policy issue (label == 1) by legislators
# based on their individual behavior mode labels (intl/ dom/ prof)

get_top_labeled_issues_by_decision <- function(df, top_n = 5) {
  # Step 1: Keep rows with valid decision labels and positive issue labels
  df_filtered <- df %>%
    filter(!is.na(decision), decision %in% c("intl", "dom", "prof"), label == 1)

   # Step 2: For each legislator, find their most frequently labeled issue (excluding congress info)
  top_issue <- df_filtered %>%
    group_by(full_name, issue) %>%
    summarise(label_count = n(), decision = first(decision), .groups = "drop") %>%
    group_by(full_name) %>%
    slice_max(label_count, n = 1, with_ties = FALSE) %>%
    ungroup()

  # Step 3: Count the most common "top issue" within each decision mode
  issue_distribution <- top_issue %>%
    group_by(decision, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(decision) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  # Step 4: Select the top N issues for each decision mode
  top_issues <- issue_distribution %>%
    group_by(decision) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(decision, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_by_label_adver <- get_top_labeled_issues_by_decision(adver_cg_tp, top_n = 3)
result_by_label_legis <- get_top_labeled_issues_by_decision(legis_cg_tp, top_n = 3)

print(result_by_label_adver$full_distribution)
print(result_by_label_legis$full_distribution)

print(result_by_label_adver$top_issues)
print(result_by_label_legis$top_issues)


# ======= function: 2. get_top_labeled_issues_by_decision_congress() =======
# this section defines a function used to count the most frequently labels issues chosen by legislators
# in each congress and decision mode 

get_top_labeled_issues_by_decision_congress <- function(df, top_n = 5) {
  
  df_filtered <- df %>%
    filter(!is.na(decision), decision %in% c("intl", "dom", "prof"), label == 1)

  top_issue <- df_filtered %>%
    group_by(full_name, congress, issue) %>%
    summarise(label_count = n(), decision = first(decision), .groups = "drop") %>%
    group_by(full_name, congress) %>%
    slice_max(label_count, n = 1, with_ties = FALSE) %>%
    ungroup()

  issue_distribution <- top_issue %>%
    group_by(congress, decision, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(congress, decision) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  top_issues <- issue_distribution %>%
    group_by(congress, decision) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(congress, decision, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_by_label_adver_congress <- get_top_labeled_issues_by_decision_congress(adver_cg_tp, top_n = 3)
result_by_label_legis_congress <- get_top_labeled_issues_by_decision_congress(legis_cg_tp, top_n = 3)


print(result_by_label_adver_congress$full_distribution)
print(result_by_label_legis_congress$full_distribution)

print(result_by_label_adver_congress$top_issues)
print(result_by_label_legis_congress$top_issues)


# ====== function: 3. get_top_labeled_issues_by_party() ======
# this section defines a function that counts the most frequently predicted issues chosen by legislators
# in each party 

get_top_labeled_issues_by_party <- function(df, top_n = 5) {
  df_filtered <- df %>%
    filter(!is.na(party), label == 1)

  top_issue <- df_filtered %>%
    group_by(full_name, issue) %>%
    summarise(label_count = n(), party = first(party), .groups = "drop") %>%
    group_by(full_name) %>%
    slice_max(label_count, n = 1, with_ties = FALSE) %>%
    ungroup()

  issue_distribution <- top_issue %>%
    group_by(party, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(party) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  top_issues <- issue_distribution %>%
    group_by(party) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(party, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

# the most emphasized issues by each party
result_by_party_adver <- get_top_labeled_issues_by_party(adver_cg_tp, top_n = 3)
result_by_party_legis <- get_top_labeled_issues_by_party(legis_cg_tp, top_n = 3)

print(result_by_party_adver$full_distribution)
print(result_by_party_legis$full_distribution)

print(result_by_party_adver$top_issues)
print(result_by_party_legis$top_issues)


# ======= function: 4. get_top_labeled_issues_by_chamber_and_decision() ========
# this section defines a function used to count the most frequently predicted issues chosen by legislators
# in each chamber and decision mode

get_top_labeled_issues_by_chamber_and_decision <- function(df, top_n = 5) {
  
  df_filtered <- df %>%
    filter(!is.na(chamber), !is.na(decision), decision %in% c("intl", "dom", "prof"), label == 1)

  top_issue <- df_filtered %>%
    group_by(full_name, issue) %>%
    summarise(label_count = n(), chamber = first(chamber), decision = first(decision), .groups = "drop") %>%
    group_by(full_name) %>%
    slice_max(label_count, n = 1, with_ties = FALSE) %>%
    ungroup()

  issue_distribution <- top_issue %>%
    group_by(chamber, decision, issue) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(chamber, decision) %>%
    arrange(desc(count)) %>%
    mutate(rank = row_number()) %>%
    ungroup()

  top_issues <- issue_distribution %>%
    group_by(chamber, decision) %>%
    slice_max(count, n = top_n, with_ties = FALSE) %>%
    arrange(chamber, decision, rank) %>%
    ungroup()

  return(list(
    full_distribution = issue_distribution,
    top_issues = top_issues
  ))
}

result_label_chamber_adver <- get_top_labeled_issues_by_chamber_and_decision(adver_cg_tp, top_n = 3)
result_label_chamber_legis <- get_top_labeled_issues_by_chamber_and_decision(legis_cg_tp, top_n = 3)

print(result_label_chamber_adver$full_distribution)
print(result_label_chamber_legis$full_distribution)

print(result_label_chamber_adver$top_issues)
print(result_label_chamber_legis$top_issues)

```

## 3.5 district_map use tigris

```{r map-preparation}
options(tigris_use_cache = TRUE)


# adver_cg_heatmap

adver_cg_heatmap <- adver_cg %>%
  filter(chamber == "House of Representatives")

# debug district data issue
unique(adver_cg_heatmap$district[is.na(as.integer(adver_cg_heatmap$district))])

# turn at-large into 00
adver_cg_heatmap  <- adver_cg_heatmap  %>%
  mutate(
    district_clean = case_when(
      grepl("At Large", district, ignore.case = TRUE) ~ "00",
      state_code %in% c("DC", "AS", "GU", "MP", "PR", "VI") ~ "98",
      grepl("^[0-9]+$", district) ~ sprintf("%02d", as.integer(district)),
      grepl("^[0-9]+(st|nd|rd|th)$", district, ignore.case = TRUE) ~ sprintf(
        "%02d", as.integer(str_extract(district, "^[0-9]+"))
      ),
      TRUE ~ NA_character_  
    ),
    state_district = paste0(state_code, "-", district_clean)
  )


unique(adver_cg_heatmap$district_clean[is.na(as.integer(adver_cg_heatmap$district_clean))]) 



# legis_cg_heatmap

legis_cg_heatmap <- legis_cg %>%
  filter(chamber == "House of Representatives")


unique(legis_cg_heatmap$district[is.na(as.integer(legis_cg_heatmap$district))])

legis_cg_heatmap  <- legis_cg_heatmap  %>%
  mutate(
    district_clean = case_when(
      grepl("At Large", district, ignore.case = TRUE) ~ "00",
      state_code %in% c("DC", "AS", "GU", "MP", "PR", "VI") ~ "98",
      grepl("^[0-9]+$", district) ~ sprintf("%02d", as.integer(district)),
      grepl("^[0-9]+(st|nd|rd|th)$", district, ignore.case = TRUE) ~ sprintf(
        "%02d", as.integer(str_extract(district, "^[0-9]+"))
      ),
      TRUE ~ NA_character_  
    ),
    state_district = paste0(state_code, "-", district_clean)
  )


unique(legis_cg_heatmap$district_clean[is.na(as.integer(legis_cg_heatmap$district_clean))]) 

# fips-state
fips_state_map <- data.frame(
  fips = c(
    "01","02","04","05","06","08","09","10","12","13","15","16","17","18","19",
    "20","21","22","23","24","25","26","27","28","29","30","31","32","33","34",
    "35","36","37","38","39","40","41","42","44","45","46","47","48","49","50",
    "51","53","54","55","56","11","60","66","69","72","78"
  ),
  state = c(
    "AL","AK","AZ","AR","CA","CO","CT","DE","FL","GA","HI","ID","IL","IN","IA",
    "KS","KY","LA","ME","MD","MA","MI","MN","MS","MO","MT","NE","NV","NH","NJ",
    "NM","NY","NC","ND","OH","OK","OR","PA","RI","SC","SD","TN","TX","UT","VT",
    "VA","WA","WV","WI","WY","DC","AS","GU","MP","PR","VI"
  ),
  stringsAsFactors = FALSE
)

# custom transformers
state_to_fips <- function(state_code) {
  res <- fips_state_map$fips[match(state_code, fips_state_map$state)]
  return(res)
}

fips_to_state <- function(fips_code) {
  # make sure fips_code is character
  fips_code <- sprintf("%02s", as.character(fips_code))
  res <- fips_state_map$state[match(fips_code, fips_state_map$fips)]
  return(res)
}

# test
fips_to_state("01")  # shoule be "AL"
fips_to_state(9)     # shoule be "CT"
fips_to_state(23)   # shoule be ME
fips_to_state(44)   # shoule be RI
fips_to_state(6)   # shoule be CA
```

```{r map-function-for-each-congress}

# ======== Function: plot_decision_map_by_congress() =========
# 


plot_decision_map_by_congress <- function(congress_num,
                                         decision_data,
                                         model_type = c("adver", "legis"),
                                         save_path = NULL,
                                         save_width = 10,
                                         save_height = 7) {
  model_type <- match.arg(model_type)
  
  # 地图年份对应
  map_year <- switch(
    as.character(congress_num),
    "115" = 2016,
    "116" = 2020,
    "117" = 2020,   # 117届用2020年选区地图
    "118" = 2022,
    stop("Unsupported congress number.")
  )
  
  # district字段对应
  district_col <- switch(
    as.character(congress_num),
    "115" = "CD115FP",
    "116" = "CD116FP",
    "117" = "CD116FP",  # 117届用CD116FP字段
    "118" = "CD118FP",
    stop("Unsupported congress number.")
  )
  
  # 载入地图数据
  cd_map <- congressional_districts(cb = TRUE, year = map_year) %>%
    mutate(
      state_code = fips_to_state(STATEFP),
      district_code = as.character(!!rlang::sym(district_col)),
      district_clean = sprintf("%02d", as.integer(district_code)),
      state_district = paste0(state_code, "-", district_clean),
      state_district = stringr::str_trim(as.character(state_district))
    )
  
  # 筛选数据
  decision_df <- decision_data %>%
    filter(congress == as.character(congress_num)) %>%
    mutate(state_district = stringr::str_trim(as.character(state_district)))
  
  # 合并地图和决策数据
  plot_df <- cd_map %>%
    left_join(decision_df %>% select(state_district, decision), by = "state_district")
  
  # 主图
  main_map <- ggplot(plot_df) +
    geom_sf(aes(fill = decision), color = "white", size = 0.1) +
    scale_fill_manual(
      values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c"),
      na.value = "grey90", name = "Decision Mode"
    ) +
    labs(
      title = paste0(
        ifelse(model_type == "adver", "Adver", "Legis"),
        " Decision Mode by District (", congress_num, "th)"
      ),
      fill = "Decision Mode"
    ) +
    theme_minimal(base_size = 14) +
    theme(legend.position = "bottom") +
    coord_sf(xlim = c(-125, -65), ylim = c(24, 50), expand = FALSE)
  
  # Alaska 小图
  alaska_map <- ggplot(subset(plot_df, state_code == "AK")) +
    geom_sf(aes(fill = decision), color = "white", size = 0.1) +
    scale_fill_manual(
      values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c"),
      na.value = "grey90", guide = "none"
    ) +
    coord_sf(xlim = c(-180, -130), ylim = c(50, 72), expand = FALSE) +
    theme_void()
  
  # Hawaii 小图
  hawaii_map <- ggplot(subset(plot_df, state_code == "HI")) +
    geom_sf(aes(fill = decision), color = "white", size = 0.1) +
    scale_fill_manual(
      values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c"),
      na.value = "grey90", guide = "none"
    ) +
    coord_sf(xlim = c(-161, -154), ylim = c(18, 23), expand = FALSE) +
    theme_void()
  
  # 合成图
  final_map <- cowplot::ggdraw() +
    cowplot::draw_plot(main_map, 0, 0, 1, 1) +
    cowplot::draw_plot(alaska_map, x = 0.03, y = 0.15, width = 0.25, height = 0.25) +
    cowplot::draw_plot(hawaii_map, x = 0.20, y = 0.15, width = 0.25, height = 0.25)
  
  if (!is.null(save_path)) {
    ggsave(filename = save_path, plot = final_map, width = save_width, height = save_height)
  }
  
  return(final_map)
}

# ========= Use of Function: plot_decision_map_by_congress()  ==========

# adver

plot_decision_map_by_congress(115, adver_cg_heatmap, model_type = "adver",
                              save_path = "adver_115_map.png")

plot_decision_map_by_congress(116, adver_cg_heatmap, model_type = "adver",
                              save_path = "adver_116_map.png")

plot_decision_map_by_congress(117, adver_cg_heatmap, model_type = "adver",
                              save_path = "adver_117_map.png")

plot_decision_map_by_congress(118, adver_cg_heatmap, model_type = "adver",
                              save_path = "adver_118_map.png")

# legis

plot_decision_map_by_congress(115, legis_cg_heatmap, model_type = "legis",
                              save_path = "legis_115_map.png")

plot_decision_map_by_congress(116, legis_cg_heatmap, model_type = "legis",
                              save_path = "legis_116_map.png")

plot_decision_map_by_congress(117, legis_cg_heatmap, model_type = "legis",
                              save_path = "legis_117_map.png")

plot_decision_map_by_congress(118, legis_cg_heatmap, model_type = "legis",
                              save_path = "legis_118_map.png")



```
### 3.5.1 analysis of geographic preferences

```{r geographic-preferences}

names(adver_cg)
names(legis_cg)
names(adver_all)

west_coast_states <- c("CA", "WA", "OR")
northeast_states <- c("ME", "NH", "VT", "MA", "RI", "CT", "NY", "NJ", "DE", "MD", "DC", "PA", "OH", "IN", "IL", "MI")
deep_red_states <- c("TX", "FL", "LA", "MS", "AL", "AR", "OK", "TN", "KY", "SC")

# 1. decision mode percentages

calculate_decision_distribution <- function(data, congress = NULL,
                                            west_coast_states = c("CA", "WA", "OR"),
                                            northeast_states = c("ME", "NH", "VT", "MA", "RI", "CT", "NY", "NJ", "DE", "MD", "DC","PA", "OH", "IN", "IL", "MI"),
                                            deep_red_states = c("TX", "FL", "LA", "MS", "AL", "AR", "OK")) {
  # 如果指定了国会届数，先过滤
  if (!is.null(congress)) {
    data <- data %>% filter(congress == !!congress)
  }
  
  # 添加地区标签
  data <- data %>%
    mutate(region = case_when(
      state_code %in% west_coast_states ~ "West Coast",
      state_code %in% northeast_states ~ "Northeast",
      state_code %in% deep_red_states ~ "Deep Red South",
      TRUE ~ "Other"
    )) %>%
    filter(region %in% c("West Coast", "Northeast", "Deep Red South"))
  
  # 计算decision模式占比
  decision_distribution <- data %>%
    group_by(region, decision) %>%
    summarise(count = n(), .groups = "drop") %>%
    group_by(region) %>%
    mutate(total = sum(count),
           proportion = count / total) %>%
    ungroup()
  
  # 添加congress列标记
  decision_distribution <- decision_distribution %>%
    mutate(congress = ifelse(is.null(congress), NA, congress))
  
  return(decision_distribution)
}


# 查看结果
# 计算adver_cg全部届数的决策模式比例
adver_result_115 <- calculate_decision_distribution(adver_cg, congress = 115)
adver_result_116 <- calculate_decision_distribution(adver_cg, congress = 116)
adver_result_117 <- calculate_decision_distribution(adver_cg, congress = 117)
adver_result_118 <- calculate_decision_distribution(adver_cg, congress = 118)

combined_adver_geographic_results <- bind_rows(
  adver_result_115,
  adver_result_116,
  adver_result_117,
  adver_result_118
)

print(combined_adver_geographic_results)


# 计算legis_cg第117届国会的决策模式比例
legis_result_115 <- calculate_decision_distribution(legis_cg, congress = 115)
legis_result_116 <- calculate_decision_distribution(legis_cg, congress = 116)
legis_result_117 <- calculate_decision_distribution(legis_cg, congress = 117)
legis_result_118 <- calculate_decision_distribution(legis_cg, congress = 118)

combined_legis_geographic_results <- bind_rows(
  legis_result_115,
  legis_result_116,
  legis_result_117,
  legis_result_118
)

print(combined_legis_geographic_results)

# 2. focal issues


# 定义函数，输入数据集和地区，输出top5议题
get_top_issues <- function(data, states_vec, top_n = 5) {
  data %>%
    filter(state_code %in% states_vec, label == 1) %>% # 过滤地区和关注标签
    group_by(issue) %>%
    summarise(count = n()) %>%
    arrange(desc(count)) %>%
    slice_head(n = top_n)
}

# 计算各地区adver_all的top5议题
adver_top_west_coast <- get_top_issues(adver_all, west_coast_states)
adver_top_northeast <- get_top_issues(adver_all, northeast_states)
adver_top_deep_red <- get_top_issues(adver_all, deep_red_states)

# 如果legis_all结构类似，计算立法议题top5
legis_top_west_coast <- get_top_issues(legis_all, west_coast_states)
legis_top_northeast <- get_top_issues(legis_all, northeast_states)
legis_top_deep_red <- get_top_issues(legis_all, deep_red_states)

# 给每个表加标识列
adver_top_west_coast <- adver_top_west_coast %>% mutate(region = "West Coast", type = "Adver")
adver_top_northeast <- adver_top_northeast %>% mutate(region = "Northeast", type = "Adver")
adver_top_deep_red <- adver_top_deep_red %>% mutate(region = "Deep Red South", type = "Adver")

legis_top_west_coast <- legis_top_west_coast %>% mutate(region = "West Coast", type = "Legis")
legis_top_northeast <- legis_top_northeast %>% mutate(region = "Northeast", type = "Legis")
legis_top_deep_red <- legis_top_deep_red %>% mutate(region = "Deep Red South", type = "Legis")

# 合并为一张表
combined_top_issues <- bind_rows(
  adver_top_west_coast,
  adver_top_northeast,
  adver_top_deep_red,
  legis_top_west_coast,
  legis_top_northeast,
  legis_top_deep_red
)

# 调整列顺序（可选）
combined_top_issues <- combined_top_issues %>%
  select(region, type, issue, count)

# 查看整合后的表格
print(combined_top_issues)

```

# 4. overall description for decision mode - each issue

## 4.1 analysis based on news_data

```{r new-data-cleansing}

news_data_path <- "/Users/kunyazhan/Desktop/American Politics/decoding us strategic competition/code/decoding code/input/115_119_news_data.csv"
news_data <- read_csv(news_data_path)


news_data_filtered <- news_data %>%
  filter(congress != 119)

news_data_aug <- news_data_filtered %>%
  mutate(congress = as.integer(congress)) %>%
  left_join(
    adver_cg %>%
      mutate(congress = as.integer(congress)) %>% 
      select(bioguide_id, congress, chamber, gender, party, committee_el),
    by = c("mid" = "bioguide_id", "congress" = "congress")
  )

news_data_filtered$congress <- as.character(news_data_filtered$congress)

missing_keys <- news_data_filtered %>%
  select(mid, congress) %>%
  distinct() %>%
  anti_join(adver_sum %>% select(bioguide_id, congress) %>% distinct(),
            by = c("mid" = "bioguide_id", "congress"))

news_data_clean <- news_data_aug %>%
  filter(!is.na(chamber), !is.na(gender), !is.na(party)) 
# already proven that these data belongs to legislators who did not continue to serve the congress

```


### 4.1.1 basic cal using original news data

```{r party-issue-attention}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- news_data_clean %>%
  group_by(congress, party, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到党派层面
issue_counts_party <- issue_counts_mid %>%
  group_by(congress, party, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每党 Top 5 议题
top_issues_party <- issue_counts_party %>%
  group_by(congress, party) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, party, desc(total_mentions))

top_issues_party

write.csv(top_issues_party, "top_issues_party_news.csv", row.names = FALSE)

```


```{r gender-issue-attention}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- news_data_clean %>%
  group_by(congress, gender, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到党派层面
issue_counts_gender <- issue_counts_mid %>%
  group_by(congress, gender, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每党 Top 5 议题
top_issues_gender <- issue_counts_gender %>%
  group_by(congress, gender) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, gender, desc(total_mentions))

top_issues_gender

write.csv(top_issues_gender, "top_issues_gender_news.csv", row.names = FALSE)

```

```{r committee-issue-attention}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- news_data_clean %>%
  group_by(congress, committee_el, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到党派层面
issue_counts_party <- issue_counts_mid %>%
  group_by(congress, committee_el, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每党 Top 5 议题
top_issues_committee_el <- issue_counts_party %>%
  group_by(congress, committee_el) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, committee_el, desc(total_mentions))

top_issues_committee_el

write.csv(top_issues_committee_el, "top_issues_party_committee.csv", row.names = FALSE)
```

```{r issue-changer-news}

issue_trend <- news_data_clean %>%
  group_by(congress, issue) %>%
  summarise(news_count = n(), .groups = "drop")


# 找出变化最剧烈的议题
issue_trend_change <- issue_trend %>%
  group_by(issue) %>%
  summarize(
    change = (news_count[congress == max(as.numeric(congress))] - news_count[congress == min(as.numeric(congress))]) /
             news_count[congress == min(as.numeric(congress))],
    .groups = "drop"
  ) %>%
  arrange(desc(abs(change)))  # 按绝对变化排序

top_issues <- issue_trend_change %>% slice(1:10) %>% pull(issue)


ggplot(issue_trend %>% filter(issue %in% top_issues),
       aes(x = as.numeric(congress), y = news_count, color = issue)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "issue change 115–118",
    x = "Congress",
    y = "news number",
    color = "issue"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Dark2") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )




# break ----

install.packages("ggbreak")
library(ggbreak)

p <- ggplot(issue_trend %>% filter(issue %in% top_issues),
       aes(x = as.numeric(congress), y = news_count, color = issue)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
     title = "Top Issues Trend Across Congress 115–118",
    x = "Congress",
    y = "news number",
    color = "issue"
  ) +
  theme_minimal() + 
  scale_color_brewer(palette = "Paired") + 
  scale_y_break(c(400, 1250)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

ggsave("issue change most 115–118_news.png", plot = p, width = 8, height = 4, dpi = 300)


# 画全部 ----

install.packages("randomcoloR")
library(randomcoloR)

n_colors <- length(unique(issue_trend$issue)) # 获取议题数量
colors <- distinctColorPalette(n_colors)

ggplot(issue_trend, aes(x = as.numeric(congress), y = news_count, color = issue)) +
  geom_line(size = 1) +
  geom_point(size = 2) +
  labs(
    title = "issue change per Congress",
    x = "Congress",
    y = "news number",
    color = "issue"
  ) +
  theme_minimal() +
  scale_color_manual(values = colors) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

p <- ggplot(issue_trend, aes(x = as.numeric(congress), y = news_count)) +
  geom_line() +
  geom_point() +
  facet_wrap(~issue, scales = "free_y") +
  labs(x = "Congress", y = "news number")


ggsave("issue change all 115–118_news.png", plot = p, width = 10, height = 8, dpi = 300)

```


```{r gender-party-issue-attention}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- news_data_clean %>%
  group_by(congress, gender, party, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到 gender × party 层面
issue_counts_gender_party <- issue_counts_mid %>%
  group_by(congress, gender, party, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每个 gender × party 的 Top 5 议题
top_issues_gender_party <- issue_counts_gender_party %>%
  group_by(congress, gender, party) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, gender, party, desc(total_mentions))

# 查看结果
top_issues_gender_party

# 导出为 CSV
write.csv(top_issues_gender_party, "top_issues_gender_party_news.csv", row.names = FALSE)
```


### 4.1.2 shanon-entropy calculation

```{r shanon-entropy-news}

# 计算每个议员在每个 congress × issue 的出现次数
issue_dist <- news_data_clean %>%
  filter(!is.na(mid), !is.na(issue), !is.na(congress)) %>%
  group_by(mid, congress, issue, gender, chamber, party) %>%
  summarise(n = n(), .groups = "drop")

# 计算每个议员在每个 congress 的总新闻数
issue_total <- issue_dist %>%
  group_by(mid, congress, gender, chamber, party) %>%
  summarise(total = sum(n), .groups = "drop")

# 合并并计算 p_k
issue_prob <- issue_dist %>%
  left_join(issue_total,
            by = c("mid", "congress", "gender", "chamber", "party")) %>%
  mutate(p = n / total)

# 计算 Shannon entropy
entropy_news <- issue_prob %>%
  group_by(mid, congress, gender, chamber, party) %>%
  summarise(
    shannon_entropy = -sum(p * log(p)),   # 默认自然对数
    n_issues = n_distinct(issue),
    total_news = sum(n),
    .groups = "drop"
  )
# 查看结果
head(entropy_news)


```


### 4.1.3 shanon-entropy pic

```{r overall-pic}

# 画图 ----

ggplot(entropy_news, aes(x = factor(congress), y = shannon_entropy)) +
  geom_violin(fill = "skyblue", alpha = 0.5) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  theme_minimal() +
  labs(x = "Congress", y = "Shannon Entropy",
       title = "Distribution of Attention Diversity per Congress")


entropy_summary <- entropy_news %>%
  group_by(congress) %>%
  summarise(mean_entropy = mean(shannon_entropy, na.rm = TRUE),
            median_entropy = median(shannon_entropy, na.rm = TRUE))
p <- ggplot(entropy_summary, aes(x = congress, y = mean_entropy)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "News Average Attention Diversity by Congress")

ggsave("Average Attention Diversity by Congress-news.png", plot = p, width = 8, height = 4, dpi = 300)

```


### 4.1.4 plus party+gender+chamber

```{r shanon-entropy-plus}

# gender
entropy_summary_gender <- entropy_news %>%
  group_by(congress, gender) %>%
  summarise(
    mean_entropy = mean(shannon_entropy, na.rm = TRUE),
    median_entropy = median(shannon_entropy, na.rm = TRUE),
    .groups = "drop"
  )

p1 <- ggplot(entropy_summary_gender, aes(x = congress, y = mean_entropy, color = gender)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "News Average Attention Diversity by Gender and Congress")


# party
entropy_summary_party <- entropy_news %>%
  group_by(congress, party) %>%
  summarise(
    mean_entropy = mean(shannon_entropy, na.rm = TRUE),
    median_entropy = median(shannon_entropy, na.rm = TRUE),
    .groups = "drop"
  )

p2 <- ggplot(entropy_summary_party, aes(x = congress, y = mean_entropy, color = party)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "News Average Attention Diversity by Party and Congress")


# chamber

 entropy_summary_chamber <- entropy_news %>%
  group_by(congress, chamber) %>%
  summarise(
    mean_entropy = mean(shannon_entropy, na.rm = TRUE),
    median_entropy = median(shannon_entropy, na.rm = TRUE),
    .groups = "drop"
  )

p3 <- ggplot(entropy_summary_chamber, aes(x = congress, y = mean_entropy, color = chamber)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "News Average Attention Diversity by Chamber and Congress")

ggsave("Average Attention Diversity by Gender and Congress-news.png", plot = p1, width = 8, height = 4, dpi = 300)
ggsave("Average Attention Diversity by Party and Congress-news.png", plot = p2, width = 8, height = 4, dpi = 300)
ggsave("Average Attention Diversity by Chamber and Congress-news.png", plot = p3, width = 8, height = 4, dpi = 300)

```


## 4.2 bills data

### 4.2.1 data cleansing

```{r shanon-entropy-bills}

bills_data_path <- "/Users/kunyazhan/Desktop/American Politics/decoding us strategic competition/code/decoding code/input/115-118_congress_data.csv"
bills_data <- read_csv(bills_data_path)


bills_data_filtered <- bills_data %>%
  filter(congress != 119)

bills_data_aug <- bills_data_filtered %>%
  mutate(congress = as.integer(congress)) %>%
  left_join(
    legis_cg %>%
      mutate(congress = as.integer(congress)) %>% 
      select(bioguide_id, congress, chamber, gender, party, committee_el),
    by = c("mid" = "bioguide_id", "congress" = "congress")
  )

summary(bills_data_aug)
sum(is.na(bills_data_aug$chamber))

```

### 4.2.2 basic cal using original bills data

```{r party-issue-attention}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- bills_data_aug %>%
  group_by(congress, party, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到党派层面
issue_counts_party <- issue_counts_mid %>%
  group_by(congress, party, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每党 Top 5 议题
top_issues_party <- issue_counts_party %>%
  group_by(congress, party) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, party, desc(total_mentions))

top_issues_party

write.csv(top_issues_party, "top_issues_party_bills.csv", row.names = FALSE)

```

```{r gender-issue-attention-bills}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- bills_data_aug %>%
  group_by(congress, gender, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到党派层面
issue_counts_gender <- issue_counts_mid %>%
  group_by(congress, gender, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每党 Top 5 议题
top_issues_gender <- issue_counts_gender %>%
  group_by(congress, gender) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, gender, desc(total_mentions))

top_issues_gender

write.csv(top_issues_gender, "top_issues_gender_bills.csv", row.names = FALSE)

```

```{r gender-party-issue-attention-bills}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- bills_data_aug %>%
  group_by(congress, gender, party, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到 gender × party 层面
issue_counts_gender_party <- issue_counts_mid %>%
  group_by(congress, gender, party, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每个 gender × party 的 Top 5 议题
top_issues_gender_party <- issue_counts_gender_party %>%
  group_by(congress, gender, party) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, gender, party, desc(total_mentions))

# 查看结果
top_issues_gender_party

# 导出为 CSV
write.csv(top_issues_gender_party, "top_issues_gender_party_bills.csv", row.names = FALSE)
```


```{r committee-issue-attention-bills}

# 1. 每个议员在每届对某议题的关注次数
issue_counts_mid <- bills_data_aug %>%
  group_by(congress, committee_el, mid, issue) %>%
  summarise(n_issue_mentions = n(), .groups = "drop")

# 2. 聚合到党派层面
issue_counts_committee_el <- issue_counts_mid %>%
  group_by(congress, committee_el, issue) %>%
  summarise(total_mentions = sum(n_issue_mentions), .groups = "drop")

# 3. 取每届每党 Top 5 议题
top_issues_committee_el <- issue_counts_committee_el %>%
  group_by(congress, committee_el) %>%
  slice_max(order_by = total_mentions, n = 5, with_ties = FALSE) %>%
  arrange(congress, committee_el, desc(total_mentions))

top_issues_committee_el

write.csv(top_issues_committee_el, "top_issues_party_committee_bills.csv", row.names = FALSE)
```

```{r issue-changer-bills}

issue_trend <- bills_data_aug %>%
  group_by(congress, issue) %>%
  summarise(news_count = n(), .groups = "drop")


# 找出变化最剧烈的议题
issue_trend_change <- issue_trend %>%
  group_by(issue) %>%
  summarize(
    change = (news_count[congress == max(as.numeric(congress))] - news_count[congress == min(as.numeric(congress))]) /
             news_count[congress == min(as.numeric(congress))],
    .groups = "drop"
  ) %>%
  arrange(desc(abs(change)))  # 按绝对变化排序

top_issues <- issue_trend_change %>% slice(1:10) %>% pull(issue)


p <- ggplot(issue_trend %>% filter(issue %in% top_issues),
       aes(x = as.numeric(congress), y = news_count, color = issue)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
    title = "Top Issues Trend Across Congress 115–118",
    x = "Congress",
    y = "bills number",
    color = "issue"
  ) +
  theme_minimal() +
  scale_color_brewer(palette = "Paired") +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

ggsave("issue change most 115–118_bills.png", plot = p, width = 8, height = 4, dpi = 300)

# break ----

ggplot(issue_trend %>% filter(issue %in% top_issues),
       aes(x = as.numeric(congress), y = news_count, color = issue)) +
  geom_line(size = 1.2) +
  geom_point(size = 2) +
  labs(
     title = "issue change 115–118",
    x = "Congress",
    y = "news number",
    color = "issue"
  ) +
  theme_minimal() + 
  scale_color_brewer(palette = "Paired") + 
  scale_y_break(c(400, 1250)) +
  theme(
    axis.text.x = element_text(angle = 45, hjust = 1),
    plot.title = element_text(hjust = 0.5)
  )

p <- ggplot(issue_trend, aes(x = as.numeric(congress), y = news_count)) +
  geom_line() +
  geom_point() +
  facet_wrap(~issue, scales = "free_y") +
  labs(x = "Congress", y = "bills number")


ggsave("issue change all 115–118_bills.png", plot = p, width = 10, height = 8, dpi = 300)


```


### 4.2.3 bills shanon entropy

```{r shanon-entropy-bills}


# 计算每个议员在每个 congress × issue 的出现次数
issue_dist <- bills_data_aug %>%
  filter(!is.na(mid), !is.na(issue), !is.na(congress)) %>%
  group_by(mid, congress, issue, gender, chamber, party) %>%
  summarise(n = n(), .groups = "drop")

# 计算每个议员在每个 congress 的总新闻数
issue_total <- issue_dist %>%
  group_by(mid, congress, gender, chamber, party) %>%
  summarise(total = sum(n), .groups = "drop")

# 合并并计算 p_k
issue_prob <- issue_dist %>%
  left_join(issue_total,
            by = c("mid", "congress", "gender", "chamber", "party")) %>%
  mutate(p = n / total)

# 计算 Shannon entropy
entropy_bills <- issue_prob %>%
  group_by(mid, congress, gender, chamber, party) %>%
  summarise(
    shannon_entropy = -sum(p * log(p)),   # 默认自然对数
    n_issues = n_distinct(issue),
    total_news = sum(n),
    .groups = "drop"
  )
# 查看结果
head(entropy_bills)

```

### 4.2.4 shanon-entropy bills pic

```{r overall-pic-bill}

# 画图 ----

ggplot(entropy_bills, aes(x = factor(congress), y = shannon_entropy)) +
  geom_violin(fill = "skyblue", alpha = 0.5) +
  geom_boxplot(width = 0.1, outlier.size = 0.5) +
  theme_minimal() +
  labs(x = "Congress", y = "Shannon Entropy",
       title = "Distribution of Attention Diversity per Congress")


entropy_summary <- entropy_bills %>%
  group_by(congress) %>%
  summarise(mean_entropy = mean(shannon_entropy, na.rm = TRUE),
            median_entropy = median(shannon_entropy, na.rm = TRUE))
p <- ggplot(entropy_summary, aes(x = congress, y = mean_entropy)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "Bills Average Attention Diversity by Congress")

ggsave("Average Attention Diversity by Congress-bills.png", plot = p, width = 8, height = 4, dpi = 300)

```


### 4.2.5 plus party+gender+chamber

```{r shanon-entropy-plus-bills}

# gender
entropy_summary_gender <- entropy_bills %>%
  group_by(congress, gender) %>%
  summarise(
    mean_entropy = mean(shannon_entropy, na.rm = TRUE),
    median_entropy = median(shannon_entropy, na.rm = TRUE),
    .groups = "drop"
  )

p1 <- ggplot(entropy_summary_gender, aes(x = congress, y = mean_entropy, color = gender)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "Bills Average Attention Diversity by Gender and Congress")


# party
entropy_summary_party <- entropy_bills %>%
  group_by(congress, party) %>%
  summarise(
    mean_entropy = mean(shannon_entropy, na.rm = TRUE),
    median_entropy = median(shannon_entropy, na.rm = TRUE),
    .groups = "drop"
  )

p2 <- ggplot(entropy_summary_party, aes(x = congress, y = mean_entropy, color = party)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "Bills Average Attention Diversity by Party and Congress")


# chamber

 entropy_summary_chamber <- entropy_bills %>%
  group_by(congress, chamber) %>%
  summarise(
    mean_entropy = mean(shannon_entropy, na.rm = TRUE),
    median_entropy = median(shannon_entropy, na.rm = TRUE),
    .groups = "drop"
  )

p3 <- ggplot(entropy_summary_chamber, aes(x = congress, y = mean_entropy, color = chamber)) +
  geom_line() +
  geom_point() +
  theme_minimal() +
  labs(x = "Congress", y = "Mean Shannon Entropy",
       title = "Bills Average Attention Diversity by Chamber and Congress")

ggsave("Average Attention Diversity by Gender and Congress-bills.png", plot = p1, width = 8, height = 4, dpi = 300)
ggsave("Average Attention Diversity by Party and Congress-bills.png", plot = p2, width = 8, height = 4, dpi = 300)
ggsave("Average Attention Diversity by Chamber and Congress-bills.png", plot = p3, width = 8, height = 4, dpi = 300)

```



## 4.3 issue analysis based on original data

```{r issue-based analysis}

# 1. issue-based dataset generation

generate_decision_issues <- function(df, set_name = "all") {
  # Step 1: Optional filter by dataset (train/test)
  if (set_name %in% c("train", "test")) {
    df <- df %>% filter(dataset == set_name)
  }

  # Step 2: Group by issue-congress level and summarize
  df_clean <- df %>%
    group_by(issue, congress, committee_el) %>%  # 保留 dataset 信息
    summarise(
      label = sum(label, na.rm = TRUE),
      pred = mean(pred, na.rm = TRUE),
      importance_intl = sum(importance_intl, na.rm = TRUE),
      importance_dom = sum(importance_dom, na.rm = TRUE),
      importance_prof = sum(importance_prof, na.rm = TRUE),
      .groups = 'drop'
    ) %>%
    mutate(
      total = importance_intl + importance_dom + importance_prof,
      intl_percentage = ifelse(total > 0, importance_intl / total, NA_real_),
      dom_percentage  = ifelse(total > 0, importance_dom  / total, NA_real_),
      prof_percentage = ifelse(total > 0, importance_prof / total, NA_real_),
      max_val = pmax(intl_percentage, dom_percentage, prof_percentage, na.rm = TRUE),
      decision = case_when(
        is.na(max_val) ~ NA_character_,
        intl_percentage == max_val & dom_percentage != max_val & prof_percentage != max_val ~ "intl",
        dom_percentage  == max_val & intl_percentage != max_val & prof_percentage != max_val ~ "dom",
        prof_percentage == max_val & intl_percentage != max_val & dom_percentage  != max_val ~ "prof",
        TRUE ~ "tie"
      )
    ) %>%
    select(-max_val, -total)

  return(df_clean)
}

# adv
adver_tp <- generate_decision_issues(adver_sum)
write.csv(adver_tp, "adver_tp.csv", row.names = FALSE)

# legis
legis_tp <- generate_decision_issues(legis_sum)
write.csv(legis_tp, "legis_tp.csv", row.names = FALSE)

unique(adver_tp$decision)
unique(legis_tp$decision)


# 2. congress heatmap

plot_decision_mode_heatmap <- function(df, title = "Decision Mode by Issue and Congress", save_path = NULL) {

  # 检查变量
  required_vars <- c("issue", "congress", "decision")
  if (!all(required_vars %in% names(df))) {
    stop("Input dataframe must contain columns: issue, congress, decision.")
  }

  # 数据准备
  df <- df %>%
    dplyr::mutate(
      decision = factor(decision, levels = c("intl", "dom", "prof")),
      congress = as.factor(congress)
    )

  # 绘图
  p <- ggplot(df, aes(x = congress, y = reorder(issue, desc(issue)), fill = decision)) +
    geom_tile(color = "white") +
    scale_fill_manual(
      values = c("intl" = "#1f77b4", "dom" = "#ff7f0e", "prof" = "#2ca02c"),
      na.value = "grey90"
    ) +
    labs(
      title = title,
      x = "Congress",
      y = "Issue",
      fill = "Decision Mode"
    ) +
    theme_minimal(base_size = 13) +
    theme(
      axis.text.x = element_text(angle = 45, hjust = 1),
      panel.grid = element_blank()
    )

  # 保存图片（如果提供了路径）
  if (!is.null(save_path)) {
    ggsave(filename = save_path, plot = p, width = 8, height = 6, dpi = 300)
  }

  return(p)
}

# use of function plot_decision_mode_heatmap()
plot_decision_mode_heatmap(legis_tp, title = "Legislative Decision Mode by Issue and Congress", save_path = "legis_mode.png")
plot_decision_mode_heatmap(adver_tp, title = "Advertise Decision Mode by Issue and Congress", save_path = "adver_mode.png")


# table
generate_decision_mode_table <- function(df) {
  # 检查变量
  required_vars <- c("issue", "congress", "decision")
  if (!all(required_vars %in% names(df))) {
    stop("Input dataframe must contain columns: issue, congress, decision.")
  }

  # 整理表格
  df_clean <- df %>%
    dplyr::mutate(
      congress = as.factor(congress),
      decision = factor(decision, levels = c("intl", "dom", "prof"))
    ) %>%
    dplyr::select(congress, issue, decision) %>%
    dplyr::arrange(congress, issue)

  return(df_clean)
}

decision_table <- generate_decision_mode_table(adver_tp)
decision_table1 <- generate_decision_mode_table(legis_tp)
print(decision_table)

write.csv(decision_table, "decision_mode_table.csv", row.names = FALSE)
write.csv(decision_table1, "decision_mode_table1.csv", row.names = FALSE)

# 3. stacked bar to show time change

legis_tp %>%
  group_by(congress, decision) %>%
  summarise(total_label = sum(label, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = factor(congress), y = total_label, fill = decision)) +
  geom_bar(stat = "identity") +
  labs(title = "Congressional Issue Attention by Decision Mode legis",
       x = "Congress", y = "Total Label") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()

adver_tp %>%
  group_by(congress, decision) %>%
  summarise(total_label = sum(label, na.rm = TRUE), .groups = "drop") %>%
  ggplot(aes(x = factor(congress), y = total_label, fill = decision)) +
  geom_bar(stat = "identity") +
  labs(title = "Congressional Issue Attention by Decision Mode legis",
       x = "Congress", y = "Total Label") +
  scale_fill_brewer(palette = "Set2") +
  theme_minimal()



# 4. congress focal issue

plot_issue_trend_by_congress <- function(df, 
                                         save_path = NULL, 
                                         file_name = "issue_trend_heatmap.png",
                                         width = 10, 
                                         height = 8, 
                                         dpi = 300,
                                         scale_type = c("raw", "normalized", "log")) {

  scale_type <- match.arg(scale_type)

  issue_congress_label <- df %>%
    filter(!is.na(issue), !is.na(congress), !is.na(label)) %>%
    group_by(congress, issue) %>%
    summarise(total_label = sum(label, na.rm = TRUE), .groups = "drop")

  # Apply scale transformation
  if (scale_type == "normalized") {
    issue_congress_label <- issue_congress_label %>%
      group_by(congress) %>%
      mutate(scaled_label = total_label / sum(total_label, na.rm = TRUE)) %>%
      ungroup()
    fill_col <- "scaled_label"
    fill_title <- "Normalized\nProportion"
  } else if (scale_type == "log") {
    issue_congress_label <- issue_congress_label %>%
      mutate(scaled_label = log1p(total_label))  # log(1 + x)
    fill_col <- "scaled_label"
    fill_title <- "Log(Label + 1)"
  } else {
    fill_col <- "total_label"
    fill_title <- "Label Count"
  }

  heatmap_plot <- ggplot(issue_congress_label, aes(x = factor(congress), y = reorder(issue, desc(issue)), fill = .data[[fill_col]])) +
    geom_tile(color = "white") +
    scale_fill_gradient(low = "#fee8c8", high = "#e34a33") +
    labs(
      title = "Most Discussed Issues by Congress",
      x = "Congress",
      y = "Issue",
      fill = fill_title
    ) +
    theme_minimal(base_size = 14) +
    theme(axis.text.x = element_text(angle = 45, hjust = 1))

  print(heatmap_plot)

  if (!is.null(save_path)) {
    ggsave(filename = save_path, plot = heatmap_plot, width = width, height = height, dpi = dpi)
  }

  return(invisible(heatmap_plot))
}

plot_issue_trend_by_congress(adver_tp, save_path = "adver_issue_trend.png", scale_type = "normalized")
plot_issue_trend_by_congress(legis_tp, save_path ="legis_issue_trend.png", scale_type = "normalized")


# table 

plot_issue_trend_by_congress <- function(df, 
                                         save_path = NULL, 
                                         file_name = "issue_trend_top5.csv",
                                         width = 10, 
                                         height = 8, 
                                         dpi = 300,
                                         scale_type = c("raw", "normalized", "log"),
                                         return_table = FALSE) {

  scale_type <- match.arg(scale_type)

  issue_congress_label <- df %>%
    filter(!is.na(issue), !is.na(congress), !is.na(label)) %>%
    group_by(congress, issue) %>%
    summarise(total_label = sum(label, na.rm = TRUE), .groups = "drop")

  # Apply scale transformation
  if (scale_type == "normalized") {
    issue_congress_label <- issue_congress_label %>%
      group_by(congress) %>%
      mutate(scaled_label = total_label / sum(total_label, na.rm = TRUE)) %>%
      ungroup()
    fill_col <- "scaled_label"
  } else if (scale_type == "log") {
    issue_congress_label <- issue_congress_label %>%
      mutate(scaled_label = log1p(total_label))  # log(1 + x)
    fill_col <- "scaled_label"
  } else {
    fill_col <- "total_label"
  }

  if (return_table) {
    # 选出每届前5个议题
    top5_table <- issue_congress_label %>%
      group_by(congress) %>%
      arrange(desc(.data[[fill_col]]), .by_group = TRUE) %>%
      slice_head(n = 5) %>%
      ungroup()

    # 如果传入了保存路径，写CSV文件
    if (!is.null(save_path)) {
      readr::write_csv(top5_table, file = save_path)
    }

    return(top5_table)

  }
}  
  
top5_adver <- plot_issue_trend_by_congress(adver_tp, scale_type = "normalized", return_table = TRUE)  
print(top5_adver)

top5_legis <- plot_issue_trend_by_congress(legis_tp, scale_type = "normalized", return_table = TRUE)  
print(top5_legis)


```

```{r issue 2}

# 5. decision mode * label * issue

ggplot(adver_tp, aes(x = factor(congress), y = reorder(issue, desc(issue)),
               size = label, color = decision)) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Issue Attention by Congress and Mode adver",
       x = "Congress", y = "Issue", size = "Label") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

legis_tp_norm <- legis_tp %>%
  group_by(congress) %>%
  mutate(label_norm = label / sum(label, na.rm = TRUE)) %>%
  ungroup()

ggplot(legis_tp_norm, aes(x = factor(congress), y = reorder(issue, desc(issue)),
               size = label_norm, color = decision)) +
  geom_point(alpha = 0.7) +
  scale_size_continuous(range = c(1, 10)) +
  labs(title = "Normalized Issue Attention by Congress and Decision Mode",
       x = "Congress", y = "Issue", size = "Normalized Label") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


ggplot(legis_tp, aes(x = factor(congress), y = reorder(issue, desc(issue)), fill = label)) +
  geom_tile(color = "white") +
  scale_fill_gradient(low = "#fee8c8", high = "#e34a33") +
  facet_wrap(~ decision, ncol = 1, scales = "free_y") +
  labs(title = "Issue Attention by Congress and Decision Mode",
       x = "Congress", y = "Issue", fill = "Label") +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# 6. sankey

install.packages("ggalluvial")
library(ggalluvial)

# 准备数据：每个 issue + decision 在不同 congress 的 label
df_alluvial <- legis_tp %>%
  filter(congress %in% c(115, 116, 117, 118)) %>%
  mutate(congress = as.factor(congress)) %>%
  group_by(issue, decision, congress) %>%
  summarise(label = sum(label, na.rm = TRUE), .groups = "drop") %>%
  arrange(issue, decision, congress)

# 确保所有列是因子类型（ggalluvial对类型敏感）
df_alluvial <- df_alluvial %>%
  mutate(
    issue = factor(issue),
    decision = factor(decision),
    congress = factor(congress)
  )

ggplot(df_alluvial,
       aes(x = congress,
           stratum = issue,
           alluvium = interaction(issue, decision),  # 不同 issue+decision 组合是不同的“流”
           y = label,
           fill = decision,
           label = issue)) +
  geom_flow(stat = "alluvium", lode.guidance = "frontback", alpha = 0.6) +
  geom_stratum(alpha = 0.8) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Issue Attention Across Congress Sessions by Decision Mode",
    x = "Congress",
    y = "Label (Attention)",
    fill = "Decision Mode"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


# with issue label

legis_issue_sankey <- ggplot(df_alluvial,
       aes(x = factor(congress), 
           stratum = issue, 
           alluvium = interaction(issue, decision),  # 不同 issue + decision 是不同“流”
           y = label,
           fill = decision,
           label = issue)) +
  geom_flow(alpha = 0.6) +
  geom_stratum(alpha = 0.8) +
  geom_text(stat = "stratum", size = 3, color = "black", min.y = 0.01) +
  scale_fill_brewer(type = "qual", palette = "Set2") +
  theme_minimal(base_size = 14) +
  labs(
    title = "Issue Attention Across Congress Sessions by Decision Mode",
    x = "Congress",
    y = "Label (Attention)",
    fill = "Decision Mode"
  ) +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))

ggsave("legis_issue_alluvial.png", plot = legis_issue_sankey, width = 14, height = 9, dpi = 300)


# 7.percentage

# 计算每届国会每种模式的议题数和占比
mode_proportion_by_congress <- adver_tp %>%
  filter(!is.na(congress), !is.na(decision), !is.na(issue)) %>%
  distinct(congress, issue, decision) %>%  # 去重，保证每个议题只计一次
  group_by(congress, decision) %>%
  summarise(issue_count = n(), .groups = "drop") %>%
  group_by(congress) %>%
  mutate(
    total_issues = sum(issue_count),
    proportion = issue_count / total_issues
  ) %>%
  ungroup()

# 查看结果
print(mode_proportion_by_congress)

```


# 5. lgbm features importance analysis 

```{r lightgbm-importance}

# selected variables for plot
importance_vars <- c("leverage_pl", "leverage_ne", 
                     "committee_pl", "committee_ne", 
                     "employ_industry_pl", "employ_industry_ne", 
                     "payroll_industry_pl", "payroll_industry_ne", 
                     "ethnicity_pl", "ethnicity_ne", 
                     "party_value_pl", "party_value_ne", 
                     "party_score_pl", "party_score_ne")

# long format
adver_long <- adver_all %>%
  select(all_of(importance_vars)) %>%
  mutate(id = row_number()) %>%
  pivot_longer(-id, names_to = "feature", values_to = "importance")

legis_long <- legis_all %>%
  select(all_of(importance_vars)) %>%
  mutate(id = row_number()) %>%
  pivot_longer(-id, names_to = "feature", values_to = "importance")


ggplot(adver_long, aes(x = importance, y = reorder(feature, importance, FUN = median))) +
  geom_violin(fill = "gray90", color = "black", alpha = 0.7) +   # 小提琴分布
  geom_jitter(height = 0.2, size = 0.4, alpha = 0.3, color = "black") +  # 样本散点
  theme_minimal(base_family = "Helvetica") +
  labs(
    x = "feature importance value (sample-level)",
    y = "feature",
    title = "adver importance of each matched features"
  ) +
  theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(size = 13, face = "bold")
  )

ggplot(legis_long, aes(x = importance, y = reorder(feature, importance, FUN = median))) +
  geom_violin(fill = "gray90", color = "black", alpha = 0.7) +   # 小提琴分布
  geom_jitter(height = 0.2, size = 0.4, alpha = 0.3, color = "black") +  # 样本散点
  theme_minimal(base_family = "Helvetica") +
  labs(
    x = "feature importance value (sample-level)",
    y = "feature",
    title = "legis importance of each matched features"
  ) +
  theme(
    axis.text.y = element_text(size = 10),
    axis.text.x = element_text(size = 10),
    plot.title = element_text(size = 13, face = "bold")
  )


# abs long format
adver_feature_sum <- adver_all %>%
  select(all_of(importance_vars)) %>%
  pivot_longer(cols = everything(), names_to = "feature", values_to = "importance") %>%
  mutate(abs_importance = abs(importance)) %>%
  group_by(feature) %>%
  summarise(total_importance = sum(abs_importance, na.rm = TRUE)) %>%
  arrange(desc(total_importance))


ggplot(adver_feature_sum, aes(x = total_importance, y = reorder(feature, total_importance))) +
  geom_col(fill = "gray20", color = "black", width = 0.6) +
  theme_minimal(base_family = "Helvetica") +
  labs(
    title = "adver importance of each matched features (abs)",
    x = "features(abs values)",
    y = "features"
  ) +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 13, face = "bold")
  )


legis_feature_sum <- legis_all %>%
  select(all_of(importance_vars)) %>%
  pivot_longer(cols = everything(), names_to = "feature", values_to = "importance") %>%
  mutate(abs_importance = abs(importance)) %>%
  group_by(feature) %>%
  summarise(total_importance = sum(abs_importance, na.rm = TRUE)) %>%
  arrange(desc(total_importance))


ggplot(legis_feature_sum, aes(x = total_importance, y = reorder(feature, total_importance))) +
  geom_col(fill = "gray20", color = "black", width = 0.6) +
  theme_minimal(base_family = "Helvetica") +
  labs(
    title = "legis importance of each matched features (abs)",
    x = "features(abs values)",
    y = "features"
  ) +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 13, face = "bold")
  )



# combine

adver_feature_sum <- adver_feature_sum %>%
  mutate(model = "adver")

legis_feature_sum <- legis_feature_sum %>%
  mutate(model = "legis")

combined_feature_sum <- bind_rows(adver_feature_sum, legis_feature_sum)

feature_order <- combined_feature_sum %>%
  group_by(feature) %>%
  summarise(total = sum(total_importance)) %>%
  arrange(desc(total)) %>%
  pull(feature)

combined_feature_sum <- combined_feature_sum %>%
  mutate(feature = factor(feature, levels = feature_order))

ggplot(combined_feature_sum, aes(x = total_importance, y = fct_rev(feature), fill = model)) +
  geom_col(position = position_dodge(width = 0.8), color = "black", width = 0.7) +
  theme_minimal(base_family = "Helvetica") +
  labs(
    title = "Absolute Feature Importance Comparison: Adver vs Legis",
    x = "Total Absolute Importance",
    y = "Feature",
    fill = "Model"
  ) +
  theme(
    axis.text = element_text(size = 10),
    plot.title = element_text(size = 14, face = "bold")
  )


```

```{r lgbm importance by congress}

# adver

# 加入 congress 后的长数据整理
adver_feature_sum_congress <- adver_all %>%
  select(congress, all_of(importance_vars)) %>%
  pivot_longer(cols = -congress, names_to = "feature", values_to = "importance") %>%
  mutate(abs_importance = abs(importance)) %>%
  group_by(congress, feature) %>%
  summarise(total_importance = sum(abs_importance, na.rm = TRUE), .groups = "drop") %>%
  arrange(congress, desc(total_importance))

# 绘图
ggplot(adver_feature_sum_congress, aes(x = total_importance, y = reorder(feature, total_importance))) +
  geom_col(fill = "gray20", color = "black", width = 0.6) +
  facet_wrap(~ congress, scales = "free_x") +
  theme_minimal(base_family = "Helvetica") +
  labs(
    title = "Importance of Each Matched Feature (by Congress, abs value)",
    x = "Total Absolute Importance",
    y = "Feature"
  ) +
  theme(
    axis.text = element_text(size = 9),
    strip.text = element_text(size = 11, face = "bold"),
    plot.title = element_text(size = 13, face = "bold")
  )


# legis

# 加入 congress 后的长数据整理
legis_feature_sum_congress <- legis_all %>%
  select(congress, all_of(importance_vars)) %>%
  pivot_longer(cols = -congress, names_to = "feature", values_to = "importance") %>%
  mutate(abs_importance = abs(importance)) %>%
  group_by(congress, feature) %>%
  summarise(total_importance = sum(abs_importance, na.rm = TRUE), .groups = "drop") %>%
  arrange(congress, desc(total_importance))

# 绘图
ggplot(legis_feature_sum_congress, aes(x = total_importance, y = reorder(feature, total_importance))) +
  geom_col(fill = "gray20", color = "black", width = 0.6) +
  facet_wrap(~ congress, scales = "free_x") +
  theme_minimal(base_family = "Helvetica") +
  labs(
    title = "Importance of Each Matched Feature (by Congress, abs value) - Legis",
    x = "Total Absolute Importance",
    y = "Feature"
  ) +
  theme(
    axis.text = element_text(size = 9),
    strip.text = element_text(size = 11, face = "bold"),
    plot.title = element_text(size = 13, face = "bold")
  )


# both

# 准备两个数据集，加入来源标签
adver_feature_sum_congress <- adver_all %>%
  select(congress, all_of(importance_vars)) %>%
  pivot_longer(cols = -congress, names_to = "feature", values_to = "importance") %>%
  mutate(abs_importance = abs(importance), source = "adver") %>%
  group_by(congress, feature, source) %>%
  summarise(total_importance = sum(abs_importance, na.rm = TRUE), .groups = "drop")

legis_feature_sum_congress <- legis_all %>%
  select(congress, all_of(importance_vars)) %>%
  pivot_longer(cols = -congress, names_to = "feature", values_to = "importance") %>%
  mutate(abs_importance = abs(importance), source = "legis") %>%
  group_by(congress, feature, source) %>%
  summarise(total_importance = sum(abs_importance, na.rm = TRUE), .groups = "drop")

# 合并两个数据框
combined_df <- bind_rows(adver_feature_sum_congress, legis_feature_sum_congress)

# 为每个congress分组内，给feature排序方便绘图

combined_df <- combined_df %>%
  group_by(congress, source) %>%
  mutate(feature_reorder = tidytext::reorder_within(feature, total_importance, congress)) %>%
  ungroup()

# 绘图
lgbm_importance_change_by_congress <- ggplot(combined_df, aes(x = total_importance, y = feature_reorder, fill = source)) +
  geom_col(position = position_dodge(width = 0.7), color = "black", width = 0.6) +
  facet_wrap(~ congress, scales = "free_y") +
  scale_y_reordered() +
  scale_fill_manual(values = c("adver" = "#1f77b4", "legis" = "#ff7f0e")) +
  theme_minimal(base_family = "Helvetica") +
  labs(
    title = "Importance of Each Matched Feature by Congress and Source",
    x = "Total Absolute Importance",
    y = "Feature",
    fill = "Source"
  ) +
  theme(
    axis.text = element_text(size = 9),
    strip.text = element_text(size = 11, face = "bold"),
    plot.title = element_text(size = 13, face = "bold")
  )

ggsave("lgbm_importance_change_by_congress.png", plot = lgbm_importance_change_by_congress, width = 14, height = 9, dpi = 300)


```


# 6. key characteristic analysis (using adver_sum/ legis_sum data)

## 6.1 database construction

```{r adver_sum cleansing}

# adv

adver_sum1 <- adver_sum %>%
  mutate(
    intl_percentage = ifelse(importance_all > 0, importance_intl / importance_all, NA_real_),
    dom_percentage  = ifelse(importance_all > 0, importance_dom / importance_all, NA_real_),
    prof_percentage = ifelse(importance_all > 0, importance_prof / importance_all, NA_real_),
    max_val = pmax(intl_percentage, dom_percentage, prof_percentage, na.rm = TRUE),
    decision = case_when(
      is.na(max_val) ~ NA_character_,
      intl_percentage == max_val & dom_percentage != max_val & prof_percentage != max_val ~ "intl",
      dom_percentage == max_val & intl_percentage != max_val & prof_percentage != max_val ~ "dom",
      prof_percentage == max_val & intl_percentage != max_val & dom_percentage != max_val ~ "prof",
      TRUE ~ "tie"
    )
  ) %>%
  select(-max_val, -importance_all)


# legis
legis_sum1 <- legis_sum %>%
  mutate(
    intl_percentage = ifelse(importance_all > 0, importance_intl / importance_all, NA_real_),
    dom_percentage  = ifelse(importance_all > 0, importance_dom / importance_all, NA_real_),
    prof_percentage = ifelse(importance_all > 0, importance_prof / importance_all, NA_real_),
    max_val = pmax(intl_percentage, dom_percentage, prof_percentage, na.rm = TRUE),
    decision = case_when(
      is.na(max_val) ~ NA_character_,
      intl_percentage == max_val & dom_percentage != max_val & prof_percentage != max_val ~ "intl",
      dom_percentage == max_val & intl_percentage != max_val & prof_percentage != max_val ~ "dom",
      prof_percentage == max_val & intl_percentage != max_val & dom_percentage != max_val ~ "prof",
      TRUE ~ "tie"
    )
  ) %>%
  select(-max_val, -importance_all)


adver_sum1 <- adver_sum1 %>%
  mutate(is.dom = ifelse(decision == "dom", 1, 0))

legis_sum1 <- legis_sum1 %>%
  mutate(is.dom = ifelse(decision == "dom", 1, 0))

```


## 6.2 adver models

### 6.2.1 regressional model 

```{r adv-digging}
# Analyze how legislators' personal characteristics affect their decision modes 


## -----------------------------
## A. regressional model
## -----------------------------
# control variables: congress

is.factor(adver_sum1$chamber)
adver_sum1$party <- as.factor(adver_sum1$party)
adver_sum1$gender <- as.factor(adver_sum1$gender)
adver_sum1$congress <- as.factor(adver_sum1$congress)
adver_sum1$chamber <- relevel(as.factor(adver_sum1$chamber), ref = "Senate")


# original 

adver_sum_intl0 <- lm(intl_percentage ~ age + 
                     party + ideology + seniority +
                     chamber + gender + 
                      as.factor(congress), 
                    data = adver_sum1)


adver_sum_dom0 <- lm(dom_percentage ~ age + 
                     party + ideology + seniority +
                     chamber + gender + 
                     as.factor(congress), 
                   data = adver_sum1)

adver_sum_prof0 <- lm(prof_percentage ~ age + 
                     party + ideology +
                     chamber + gender + 
                      as.factor(congress), 
                    data = adver_sum1)

# output regression summaries 
summary(adver_sum_intl0)
summary(adver_sum_dom0)
summary(adver_sum_prof0)


models1 <- list(
  "International" = adver_sum_intl0,
  "Domestic"      = adver_sum_dom0,
  "Professional"  = adver_sum_prof0
)

modelsummary(
  models1,
  output = "original_model_result_adver.html",
  stars = TRUE,
  statistic = "std.error"
)




# check basic assumptions of linear regression
qqnorm(residuals(adver_sum_prof0))
qqline(residuals(adver_sum_prof0), col = "red")
mean(residuals(adver_sum_prof0))
vif(adver_sum_prof0)

qqnorm(residuals(adver_sum_dom0))
qqline(residuals(adver_sum_dom0), col = "red")
mean(residuals(adver_sum_dom0))
vif(adver_sum_dom0)


qqnorm(residuals(adver_sum_intl0))
qqline(residuals(adver_sum_intl0), col = "red")
mean(residuals(adver_sum_intl0))
vif(adver_sum_intl0)


# exact observations used
nrow(model.frame(adver_sum_prof0))  
nrow(model.frame(adver_sum_intl0))  
nrow(model.frame(adver_sum_dom0))  
```

### 6.2.2 ridge & lasso 

```{r adver_sum ridge&lasso}
## -----------------------------
## b. Ridge & lasso
## -----------------------------

# Step 1: Clean and filter complete data
adver_sum_clean <- adver_sum1 %>%
  filter(
    !is.na(intl_percentage),
    !is.na(dom_percentage),
    !is.na(prof_percentage),
    !is.na(age),
    !is.na(party),
    !is.na(seniority),
    !is.na(ideology),
    !is.na(chamber),
    !is.na(gender),
    !is.na(congress)
  )

# Step 2: constructing matrix, and use adver_sum_clean
x_clean <- model.matrix(~ age + party + seniority + ideology + chamber + gender + as.factor(congress), data = adver_sum_clean)[, -1]

# Step 3: dependent variables
y_intl <- adver_sum_clean$intl_percentage
y_dom  <- adver_sum_clean$dom_percentage
y_prof <- adver_sum_clean$prof_percentage


# Ridge（alpha=0）
cv_ridge_intl <- cv.glmnet(x_clean, y_intl, alpha=0)
best_lambda_intl <- cv_ridge_intl$lambda.min
ridge_model_intl <- glmnet(x_clean, y_intl, alpha=0, lambda=best_lambda_intl)

cv_ridge_dom <- cv.glmnet(x_clean, y_dom, alpha=0)
best_lambda_dom <- cv_ridge_dom$lambda.min
ridge_model_dom <- glmnet(x_clean, y_dom, alpha=0, lambda=best_lambda_dom)

cv_ridge_prof <- cv.glmnet(x_clean, y_prof, alpha=0)
best_lambda_prof <- cv_ridge_prof$lambda.min
ridge_model_prof <- glmnet(x_clean, y_prof, alpha=0, lambda=best_lambda_prof)

# Lasso（alpha=1）
cv_lasso_intl <- cv.glmnet(x_clean, y_intl, alpha=1)
best_lambda_lasso_intl <- cv_lasso_intl$lambda.min
lasso_model_intl <- glmnet(x_clean, y_intl, alpha=1, lambda=best_lambda_lasso_intl)

cv_lasso_dom <- cv.glmnet(x_clean, y_dom, alpha=1)
best_lambda_lasso_dom <- cv_lasso_dom$lambda.min
lasso_model_dom <- glmnet(x_clean, y_dom, alpha=1, lambda=best_lambda_lasso_dom)

cv_lasso_prof <- cv.glmnet(x_clean, y_prof, alpha=1)
best_lambda_lasso_prof <- cv_lasso_prof$lambda.min
lasso_model_prof <- glmnet(x_clean, y_prof, alpha=1, lambda=best_lambda_lasso_prof)

# Output model coefficients

coef(ridge_model_intl)
coef(ridge_model_dom)
coef(ridge_model_prof)

coef(lasso_model_intl)
coef(lasso_model_dom)
coef(lasso_model_prof)

# Output tidy tables
get_tidy_coefs <- function(model, mode, method) {
  tidy(model) %>%
    mutate(mode = mode, method = method)
}

# combine all models
all_models <- bind_rows(
  get_tidy_coefs(ridge_model_intl, "intl", "ridge"),
  get_tidy_coefs(ridge_model_dom, "dom", "ridge"),
  get_tidy_coefs(ridge_model_prof, "prof", "ridge"),
  get_tidy_coefs(lasso_model_intl, "intl", "lasso"),
  get_tidy_coefs(lasso_model_dom, "dom", "lasso"),
  get_tidy_coefs(lasso_model_prof, "prof", "lasso")
)

# 确保先加载 dplyr 和 tidyr
library(dplyr)
library(tidyr)
library(knitr)
library(kableExtra)

# 强制使用 dplyr::select 和 tidyr::pivot_wider 避免冲突
coef_table <- all_models %>%
  dplyr::select(term, estimate, mode, method) %>%
  dplyr::mutate(model_label = paste0(method, "_", mode)) %>%
  dplyr::select(term, estimate, model_label) %>%
  tidyr::pivot_wider(names_from = model_label, values_from = estimate)

# 构建 HTML 表格
html_table <- coef_table %>%
  kable(format = "html", digits = 4) %>%
  kable_styling(full_width = FALSE)

# 保存为 HTML 文件
writeLines(as.character(html_table), "coef_table.html")
```


## 6.2 legis model

### 6.2.1 regressional model

```{r legis-digging}

## -----------------------------
# a. regressional model
## -----------------------------

legis_sum1$party <- as.factor(legis_sum1$party)
legis_sum1$gender <- as.factor(legis_sum1$gender)
legis_sum1$congress <- as.factor(legis_sum1$congress)
legis_sum1$chamber <- relevel(as.factor(legis_sum1$chamber), ref = "Senate")


# original 

legis_sum_intl0 <- lm(intl_percentage ~ age + 
                     party + ideology + seniority+ 
                     chamber + gender + 
                      as.factor(congress), 
                    data = legis_sum1)


legis_sum_dom0 <- lm(dom_percentage ~ age + 
                     party + ideology + seniority+ 
                     chamber + gender + 
                     as.factor(congress), 
                   data = legis_sum1)

legis_sum_prof0 <- lm(prof_percentage ~ age + 
                     party + ideology + seniority+ 
                     chamber + gender + 
                      as.factor(congress), 
                    data = legis_sum1)

# output regression summaries 
summary(legis_sum_intl0)
summary(legis_sum_dom0)
summary(legis_sum_prof0)


models_legis1 <- list(
  "International" = legis_sum_intl0,
  "Domestic"      = legis_sum_dom0,
  "Professional"  = legis_sum_prof0
)

modelsummary(
  models_legis1,
  output = "original_model_result_legis.html",
  stars = TRUE,
  statistic = "std.error"
)


# adjusted according to lasso and ridge


legis_sum_intl <- lm(intl_percentage ~ age + 
                      party + seniority +
                      chamber + gender 
                    + as.factor(congress),   data = legis_sum1)

legis_sum_dom <- lm(dom_percentage ~ age + 
                     party + 
                     chamber + gender 
                     + as.factor(congress),   data = legis_sum1)

legis_sum_prof <- lm(prof_percentage ~ age + 
                       chamber + ideology 
                      + as.factor(congress),   data = legis_sum1)

# assumption check
qqnorm(residuals(legis_sum_intl))
qqline(residuals(legis_sum_intl), col = "red")
mean(residuals(legis_sum_intl))
vif(legis_sum_intl)


qqnorm(residuals(legis_sum_dom))
qqline(residuals(legis_sum_dom), col = "red")
mean(residuals(legis_sum_dom))
vif(legis_sum_dom)


qqnorm(residuals(legis_sum_prof))
qqline(residuals(legis_sum_prof), col = "red")
mean(residuals(legis_sum_prof))
vif(legis_sum_prof)


# results of model a
summary(legis_sum_intl)
summary(legis_sum_dom)
summary(legis_sum_prof)


# output
models_legis2 <- list(
  "International" = legis_sum_intl,
  "Domestic"      = legis_sum_dom,
  "Professional"  = legis_sum_prof
)

modelsummary(
  models_legis2,
  output = "adjusted_model_result_legis.html",
  stars = TRUE,
  statistic = "std.error"
)
```


```{r legis_sum ridge&lasso}
## -----------------------------
## c. ridge & lasso 
## -----------------------------

legis_sum_clean <- legis_sum1 %>%
  filter(
    !is.na(intl_percentage),
    !is.na(age),
    !is.na(party),
    !is.na(seniority),
    !is.na(ideology),
    !is.na(chamber),
    !is.na(gender),
    !is.na(congress)
  )

# matrix and dependent variable
x_clean <- model.matrix(~ age + party + seniority + ideology + chamber + gender + as.factor(congress), data=legis_sum_clean)[, -1]
y_intl <- legis_sum_clean$intl_percentage
y_dom <- legis_sum_clean$dom_percentage
y_prof <- legis_sum_clean$prof_percentage

# Ridge（alpha=0）
cv_ridge_intl <- cv.glmnet(x_clean, y_intl, alpha=0)
best_lambda_intl <- cv_ridge_intl$lambda.min
ridge_model_intl <- glmnet(x_clean, y_intl, alpha=0, lambda=best_lambda_intl)

cv_ridge_dom <- cv.glmnet(x_clean, y_dom, alpha=0)
best_lambda_dom <- cv_ridge_dom$lambda.min
ridge_model_dom <- glmnet(x_clean, y_dom, alpha=0, lambda=best_lambda_dom)

cv_ridge_prof <- cv.glmnet(x_clean, y_prof, alpha=0)
best_lambda_prof <- cv_ridge_prof$lambda.min
ridge_model_prof <- glmnet(x_clean, y_prof, alpha=0, lambda=best_lambda_prof)

# Lasso（alpha=1）
cv_lasso_intl <- cv.glmnet(x_clean, y_intl, alpha=1)
best_lambda_lasso_intl <- cv_lasso_intl$lambda.min
lasso_model_intl <- glmnet(x_clean, y_intl, alpha=1, lambda=best_lambda_lasso_intl)

cv_lasso_dom <- cv.glmnet(x_clean, y_dom, alpha=1)
best_lambda_lasso_dom <- cv_lasso_dom$lambda.min
lasso_model_dom <- glmnet(x_clean, y_dom, alpha=1, lambda=best_lambda_lasso_dom)

cv_lasso_prof <- cv.glmnet(x_clean, y_prof, alpha=1)
best_lambda_lasso_prof <- cv_lasso_prof$lambda.min
lasso_model_prof <- glmnet(x_clean, y_prof, alpha=1, lambda=best_lambda_lasso_prof)


coef(ridge_model_intl)
coef(ridge_model_dom)
coef(ridge_model_prof)

coef(lasso_model_intl)
coef(lasso_model_dom)
coef(lasso_model_prof)

# output
get_tidy_coefs <- function(model, mode, method) {
  tidy(model) %>%
    mutate(mode = mode, method = method)
}

all_models <- bind_rows(
  get_tidy_coefs(ridge_model_intl, "intl", "ridge"),
  get_tidy_coefs(ridge_model_dom, "dom", "ridge"),
  get_tidy_coefs(ridge_model_prof, "prof", "ridge"),
  get_tidy_coefs(lasso_model_intl, "intl", "lasso"),
  get_tidy_coefs(lasso_model_dom, "dom", "lasso"),
  get_tidy_coefs(lasso_model_prof, "prof", "lasso")
)



# 强制使用 dplyr::select 和 tidyr::pivot_wider 避免冲突
coef_table1 <- all_models %>%
  dplyr::select(term, estimate, mode, method) %>%
  dplyr::mutate(model_label = paste0(method, "_", mode)) %>%
  dplyr::select(term, estimate, model_label) %>%
  tidyr::pivot_wider(names_from = model_label, values_from = estimate)

# 构建 HTML 表格
html_table1 <- coef_table1 %>%
  kable(format = "html", digits = 4) %>%
  kable_styling(full_width = FALSE)

# 保存为 HTML 文件
writeLines(as.character(html_table1), "coef_table1.html")

```


# 7. dependent variable changed into is.x and x_percentage

## 7.1 adver/legis_cg cleansing


```{r analysis-data-construction}


adver_cg <- generate_decision_with_traits(adver_sum)
legis_cg <- generate_decision_with_traits(legis_sum)


# if y = is.dom
adver_cg <- adver_cg %>%
  mutate(is.dom = ifelse(decision == "dom", 1, 0))
legis_cg <- legis_cg %>%
  mutate(is.dom = ifelse(decision == "dom", 1, 0))

# if y = is.prof
adver_cg <- adver_cg %>%
  mutate(is.prof = ifelse(decision == "prof", 1, 0))
legis_cg <- legis_cg %>%
  mutate(is.prof = ifelse(decision == "prof", 1, 0))

# not applicable for intl, cuz there is no "intl" in adver


adver_cg <- adver_cg %>% 
  mutate(is.republican = ifelse(party == "Republican", 1, 0))

legis_cg <- legis_cg %>% 
  mutate(is.republican = ifelse(party == "Republican", 1, 0))


# turn categorical vector into numeric 
adver_cg$gender_numeric <- ifelse(adver_cg$gender == "Male", 1, 0)
legis_cg$gender_numeric <- ifelse(legis_cg$gender == "Male", 1, 0)

legis_cg$chamber_numeric <- ifelse(legis_cg$chamber == "Senate", 1, 0)
adver_cg$chamber_numeric <- ifelse(adver_cg$chamber == "Senate", 1, 0)


# 0. variables colinearity 

cor(adver_cg$age, adver_cg$seniority, use = "complete.obs")
cor(legis_cg$age, legis_cg$seniority, use = "complete.obs")

boxplot(ideology ~ is.republican, data = adver_cg)
boxplot(ideology ~ is.republican, data = legis_cg)
cor.test(adver_cg$ideology, adver_cg$is.republican, method = "pearson") 

boxplot(ideology ~ chamber, data = adver_cg)
boxplot(ideology ~ chamber, data = legis_cg)
cor.test(adver_cg$ideology, adver_cg$chamber_numeric, method = "pearson") 

boxplot(ideology ~ gender, data = adver_cg)
boxplot(ideology ~ gender, data = legis_cg)
cor.test(adver_cg$ideology, adver_cg$gender_numeric, method = "pearson") 



# is.republican vs chamber
table(adver_cg$is.republican, adver_cg$chamber)
chisq.test(table(adver_cg$is.republican, adver_cg$chamber))

table(legis_cg$is.republican, legis_cg$chamber)
chisq.test(table(legis_cg$is.republican, legis_cg$chamber))


# is.republican vs gender
table(adver_cg$is.republican, adver_cg$gender)
chisq.test(table(adver_cg$is.republican, adver_cg$chamber))

table(legis_cg$is.republican, legis_cg$chamber)
chisq.test(table(legis_cg$is.republican, legis_cg$chamber))

```

## 7.2 residual test 

```{r residual test for each model}

install.packages("ResourceSelection")
library(ResourceSelection)

length(adver_cg$is.dom)
length(fitted(adver_logit0))

# lm




# logit
hoslem.test(model.frame(adver_logit1)$is.dom, fitted(adver_logit1))

model_data <- model.frame(adver_logit1)

y <- model.response(model_data)
y <- as.numeric(as.character(y))

yhat <- fitted(adver_logit1)
length(y); length(yhat)

hoslem.test(y, yhat, g = 10)





```


## 7.3 is.dom 

```{r logit-models-construction}

## -----------------------------
# a. logit model - multiple variables
## -----------------------------

# 0. lasso

run_glmnet_selection <- function(data, formula, alpha = 1, family = "binomial") {
  # 处理缺失值
  complete_data <- model.frame(formula, data = data, na.action = na.omit)
  
  # 自变量矩阵
  X <- model.matrix(formula, data = complete_data)[,-1]
  
  # 因变量
  y <- complete_data[[as.character(formula[[2]])]]
  
  # 拟合 lasso/ridge/elnet
  cv_fit <- cv.glmnet(X, y, family = family, alpha = alpha)
  
  # 可视化交叉验证误差
  plot(cv_fit, main = paste("CV for alpha =", alpha, ", family =", family))
  
  # 输出最佳 lambda 和系数
  cat("Best lambda:", cv_fit$lambda.min, "\n")
  print(coef(cv_fit, s = "lambda.min"))
  
  return(cv_fit)
}


# adver
cv_lasso_adver <- run_glmnet_selection(adver_cg, is.dom ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 1,
  family = "binomial")
cv_ridge_adver <- run_glmnet_selection(adver_cg, is.dom ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0,
  family = "binomial")
cv_elnet_adver <- run_glmnet_selection(adver_cg, is.dom ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0.5,
  family = "binomial")

# legis
cv_lasso_legis <- run_glmnet_selection(legis_cg, is.dom ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 1,
  family = "binomial")
cv_ridge_legis <- run_glmnet_selection(legis_cg, is.dom ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0,
  family = "binomial")
cv_elnet_legis <- run_glmnet_selection(legis_cg, is.dom ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0.5,
  family = "binomial")


# 1.  model construction

analyze_logit_model <- function(model) {
  library(broom)
  
  tidy_model <- tryCatch(
    tidy(model, conf.int = TRUE),
    error = function(e) {
      message("Error in tidy(): ", e$message)
      return(NULL)
    }
  )
  
  if (is.null(tidy_model)) return(NULL)
  
  # 检查是否包含 conf.low 和 conf.high，若无则用 NA 填充
  if (!"conf.low" %in% names(tidy_model)) tidy_model$conf.low <- NA
  if (!"conf.high" %in% names(tidy_model)) tidy_model$conf.high <- NA
  
  coefs <- tidy_model %>%
    mutate(
      odds_ratio = exp(estimate),
      conf.low.or = exp(conf.low),
      conf.high.or = exp(conf.high),
      significance = case_when(
        p.value < 0.001 ~ "***",
        p.value < 0.01  ~ "**",
        p.value < 0.05  ~ "*",
        p.value < 0.1   ~ ".",
        TRUE ~ ""
      )
    ) %>%
    select(term, estimate, odds_ratio, conf.low.or, conf.high.or,
           std.error, p.value, significance)
  
  print(summary(model))
  print(coefs)
  
  # pseudo R-squared (McFadden)
  if (requireNamespace("pscl", quietly = TRUE)) {
    library(pscl)
    cat("\nPseudo R-squared:\n")
    print(pscl::pR2(model))
  } else {
    message("Install 'pscl' package for pseudo R-squared calculation.")
  }
  
  return(coefs)
}

# 2. set references

set_reference_levels <- function(df,
                                 ref_party = "democratic",
                                 ref_is.republican = "0",
                                 ref_chamber = "Senate") {
  
  if ("party" %in% names(df)) {
    df$party <- factor(df$party)  # 转换为factor
    df$party <- relevel(df$party, ref = ref_party)  # 设置reference level
  }
  
   if ("party" %in% names(df)) {
    df$is.republican <- factor(df$is.republican)  # 转换为factor
    df$is.republican <- relevel(df$is.republican, ref = ref_is.republican)  # 设置reference level
  }
  
  if ("chamber" %in% names(df)) {
    df$chamber <- factor(df$chamber)
    df$chamber <- relevel(df$chamber, ref = ref_chamber)
  }
  
  return(df)
}


adver_cg <- set_reference_levels(adver_cg,
                                 ref_party = "Democratic",
                                 ref_is.republican = "0",
                                 ref_chamber = "Senate")

legis_cg <- set_reference_levels(legis_cg,
                                 ref_party = "Democratic",
                                 ref_is.republican = "0",
                                 ref_chamber = "Senate")

levels(adver_cg$chamber)
levels(adver_cg$party)
levels(adver_cg$is.republican)

# 3. model trainning
#adver
adver_logit0 <- glm(is.dom ~ age + is.republican + ideology + chamber + seniority + gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())


adver_logit_result <- analyze_logit_model(adver_logit0)


#legis
legis_logit0 <- glm(is.dom ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_result <- analyze_logit_model(legis_logit0)


# ajusted

adver_logit1 <- glm(is.dom ~ age +  chamber + ideology +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())
adver_logit_result <- analyze_logit_model(adver_logit1)

legis_logit1 <- glm(is.dom ~ age +  chamber +ideology +  gender+as.factor(congress),  
                   data = legis_cg, 
                   family = binomial())
legis_logit_result <- analyze_logit_model(legis_logit1)


adver_logit2 <- glm(is.dom ~ age +  chamber + is.republican +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())
adver_logit_result <- analyze_logit_model(adver_logit2)

legis_logit2 <- glm(is.dom ~ age +  chamber + is.republican +  gender+as.factor(congress),  
                   data = legis_cg, 
                 family = binomial())
legis_logit_result <- analyze_logit_model(legis_logit2)  

adver_logit3 <- glm(is.dom ~ seniority +  chamber + is.republican +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

legis_logit3 <- glm(is.dom ~ seniority +  chamber + is.republican +  gender+as.factor(congress),  
                   data = legis_cg, 
                   family = binomial())


adver_logit4 <- glm(is.dom ~ seniority +  chamber + ideology +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

legis_logit4 <- glm(is.dom ~ seniority +  chamber + ideology +  gender+as.factor(congress),  
                   data = legis_cg, 
                   family = binomial())



# 4. check 

# pesudo r2
pR2(adver_logit1)
pR2(legis_logit1)
pR2(adver_logit2)
pR2(legis_logit2)
pR2(adver_logit3)
pR2(legis_logit3)
pR2(adver_logit4)
pR2(legis_logit4)


vif(adver_logit1)
vif(legis_logit1)
vif(adver_logit2)
vif(legis_logit2)
vif(adver_logit3)
vif(legis_logit3)
vif(adver_logit4)
vif(legis_logit4)


# robust check on legis_logit


# 5. output

modelsummary(
  list(
    "Adver All" = adver_logit0,
    "1" = adver_logit1,
    "2" = adver_logit2,
    "3" = adver_logit3,
    "4" = adver_logit4
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_adver_isdom.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)

modelsummary(
  list(
    "Legis All" = legis_logit0,
    "1" = legis_logit1,
    "2" = legis_logit2,
    "3" = legis_logit3,
    "4" = legis_logit4
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_legis_isdom.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)


# coefficient plot / dot-whisker plot

# 整理结果
models <- list(Age = adver_logit_age, Ideology = adver_logit_ideology, Party = adver_logit_is.republican, Gender = adver_logit_gender, Chamber = adver_logit_chamber, Seniority = adver_logit_seniority)
tidy_models <- purrr::map_df(models, tidy, .id = "Model")

# 画图
p <- gp <- ggplot(tidy_models, aes(x = estimate, y = term, color = Model)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = estimate - 1.96*std.error,
                     xmax = estimate + 1.96*std.error),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Logit Coefficient", y = "Variable", title = "Coefficient Plot for Adver Logit seperated regression") 

ggsave("adver_seperated.png", plot = p, width = 8, height = 4, dpi = 300)

## -----------------------------
# b. logit model - multiple variables
## -----------------------------

# 1.1 logit model construction

#adver
adver_logit_age <- glm(is.dom ~ age + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_ideology <- glm(is.dom ~ ideology + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_is.republican <- glm(is.dom ~ is.republican + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_gender <- glm(is.dom ~ gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_chamber <- glm(is.dom ~ chamber + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_seniority <- glm(is.dom ~ seniority + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())


#legis
legis_logit_age <- glm(is.dom ~ age + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_ideology <- glm(is.dom ~ ideology + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_is.republican <- glm(is.dom ~ is.republican + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_gender <- glm(is.dom ~ gender + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_chamber <- glm(is.dom ~ chamber + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_seniority <- glm(is.dom ~ seniority + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())


# 1.2. output

# adver
modelsummary(
  list(
    "1" = adver_logit_age,
    "2" = adver_logit_ideology,
    "3" = adver_logit_is.republican,
    "4" = adver_logit_gender,
    "5" = adver_logit_chamber,
    "6" = adver_logit_seniority
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_adver_seperated_isdom.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)


#legis
modelsummary(
  list(
    "1" = legis_logit_age,
    "2" = legis_logit_ideology,
    "3" = legis_logit_is.republican,
    "4" = legis_logit_gender,
    "5" = legis_logit_chamber,
    "6" = legis_logit_seniority
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_legis_seperated_isdom.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)


# coefficient plot / dot-whisker plot

# 整理结果
models <- list(Age = legis_logit_age, Ideology = legis_logit_ideology, Party = legis_logit_is.republican, Gender = legis_logit_gender, Chamber = legis_logit_chamber, Seniority = legis_logit_seniority)
tidy_models <- purrr::map_df(models, tidy, .id = "Model")

# 画图
p <- gp <- ggplot(tidy_models, aes(x = estimate, y = term, color = Model)) +
  geom_point(position = position_dodge(width = 0.5)) +
  geom_errorbarh(aes(xmin = estimate - 1.96*std.error,
                     xmax = estimate + 1.96*std.error),
                 height = 0.2,
                 position = position_dodge(width = 0.5)) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  theme_minimal() +
  labs(x = "Logit Coefficient", y = "Variable", title = "Coefficient Plot for Legis Logit seperated regression") 

ggsave("legis_seperated.png", plot = p, width = 8, height = 4, dpi = 300)


# 1.3. check ----

# pesudo r2
pR2(adver_logit_age)
pR2(adver_logit_ideology)
pR2(adver_logit_is.republican)
pR2(adver_logit_gender)
pR2(adver_logit_seniority)
pR2(adver_logit_chamber)

pR2(legis_logit_age)
pR2(legis_logit_ideology)
pR2(legis_logit_is.republican)
pR2(legis_logit_gender)
pR2(legis_logit_seniority)
pR2(legis_logit_chamber)


vif(adver_logit_age)
vif(adver_logit_ideology)
vif(adver_logit_is.republican)
vif(adver_logit_gender)
vif(adver_logit_seniority)
vif(adver_logit_chamber)

vif(legis_logit_age)
vif(legis_logit_ideology)
vif(legis_logit_is.republican)
vif(legis_logit_gender)
vif(legis_logit_seniority)
vif(legis_logit_chamber)

# residual

library(DHARMa)

plot_logit_residuals <- function(model_list, data, output_dir = "logit_residual_plots", n = 1000) {
  
  # 如果输出文件夹不存在，就新建
  if (!dir.exists(output_dir)) {
    dir.create(output_dir)
  }
  
  for (model_name in names(model_list)) {
    model <- model_list[[model_name]]
    
    # 生成残差
    res <- simulateResiduals(fittedModel = model, n = n)
    
    # 自动获取自变量（去掉因变量）
    predictors <- all.vars(formula(model))[-1]
    predictors <- predictors[predictors %in% colnames(data)]  # 只保留存在的列
    
    # 设置输出文件（一个大图）
    png(file.path(output_dir, paste0("residuals_", model_name, "_all.png")),
        width = 2400, height = 1600, res = 200)
    
    # 设置大图布局：根据变量个数自动排版（2列）
    n_pred <- length(predictors)
    n_row <- ceiling(n_pred / 2)
    par(mfrow = c(n_row, 2), mar = c(4, 4, 3, 1))  # 调整边距，避免标题重叠
    
    # 遍历 predictor 画子图
    for (var in predictors) {
      plotResiduals(res, predictor = data[[var]],
                    main = paste("Residuals vs", var, "\n(", model_name, ")"))
    }
    
    dev.off()
  }
  
  cat("✅ 每个模型的所有 predictor 残差图已合并保存至：", output_dir, "\n")
}

# ----------------------------

# 模型列表
model_list_adver_dom_logit <- list(
  "adver_dom_logit1" = adver_logit1,
  "adver_dom_logit2" = adver_logit2,
  "adver_dom_logit3" = adver_logit3,
  "adver_dom_logit4" = adver_logit4
)

model_list_legis_dom_logit <- list(
  "legis_dom_logit1" = legis_logit1,
  "legis_dom_logit2" = legis_logit2,
  "legis_dom_logit3" = legis_logit3,
  "legis_dom_logit4" = legis_logit4
)

# 调用函数
plot_logit_residuals(model_list_adver_dom_logit, adver_cg)
plot_logit_residuals(model_list_legis_dom_logit, legis_cg)






```


## 7.4 dom_percentage

```{r lm-models-construction}


## -----------------------------
# a. regression model - mutilple variables
## -----------------------------

# 1. lasso/ ridge ----

# adver
cv_lasso_adver1 <- run_glmnet_selection(
  data = adver_cg,
  formula = dom_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 1,
  family = "gaussian"  
)

cv_ridge_adver1 <- run_glmnet_selection(
  data = adver_cg,
  formula = dom_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0,
  family = "gaussian"  
)

cv_elnet_adver1 <- run_glmnet_selection(
  data = adver_cg,
  formula = dom_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0.5,
  family = "gaussian"  
)

# legis
cv_lasso_legis1 <- run_glmnet_selection(
  data = legis_cg,
  formula = dom_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 1,
  family = "gaussian"  
)

cv_ridge_legis1 <- run_glmnet_selection(
  data = legis_cg,
  formula = dom_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0,
  family = "gaussian"  
)

cv_elnet_legis1 <- run_glmnet_selection(
  data = legis_cg,
  formula = dom_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0.5,
  family = "gaussian"  
)


# 2. model training ----
# original
adver_lm0 <- lm(dom_percentage ~ age + 
                     is.republican + ideology + seniority+ 
                     chamber + gender + 
                     as.factor(congress), 
                   data = adver_cg)

legis_lm0 <- lm(dom_percentage ~ age + 
                     is.republican + ideology + seniority+ 
                     chamber + gender + 
                     as.factor(congress), 
                   data = legis_cg)

summary(adver_lm0)
summary(legis_lm0)

# adjusted

adver_lm1 <- lm(dom_percentage ~ age + 
                     is.republican + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


vif(lm(dom_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress), data = adver_cg))
vif(lm(dom_percentage ~ age + is.republican + ideology + chamber + gender + as.factor(congress), data = adver_cg))
vif(lm(dom_percentage ~  seniority + is.republican + ideology + chamber + gender + as.factor(congress), data = adver_cg))



legis_lm1 <- lm(dom_percentage ~ age + 
                     chamber + gender + is.republican 
                     + as.factor(congress),   data = legis_cg )


adver_lm2 <- lm(dom_percentage ~ age + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


legis_lm2 <- lm(dom_percentage ~ age + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = legis_cg)


adver_lm3 <- lm(dom_percentage ~ seniority + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


legis_lm3 <- lm(dom_percentage ~ seniority + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = legis_cg)


adver_lm4 <- lm(dom_percentage ~ seniority + 
                     is.republican + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


legis_lm4 <- lm(dom_percentage ~ seniority + 
                     is.republican + 
                     chamber + gender +
                     as.factor(congress), 
                   data = legis_cg)

# 3. qqplot check

# check basic assumptions of linear regression
#qqnorm(residuals(adver_cg_dom))
#qqline(residuals(adver_cg_dom), col = "red")
#mean(residuals(adver_cg_dom))
#vif(adver_cg_dom)

#qqnorm(residuals(legis_cg_dom))
#qqline(residuals(legis_cg_dom), col = "red")
#mean(residuals(legis_cg_dom))
#vif(legis_cg_dom)






# 4. output

modelsummary(
  list(
    "Adver LM" = adver_lm0,
    "1" = adver_lm1,
    "2" = adver_lm2,
    "3" = adver_lm3,
    "4" = adver_lm4
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodels_adver_dom.html",
  escape = FALSE
)

modelsummary(
  list(
    "Legis LM" = legis_lm0,
    "1" = legis_lm1,
    "2" = legis_lm2,
    "3" = legis_lm3,
    "4" = legis_lm4
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodels_legis_dom.html",
  escape = FALSE
)


## -----------------------------
# a. regression model - seperated variables
## -----------------------------


# 2.1. lm model construction

# adver
adver_cg_age1 <- lm(dom_percentage ~ age +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_ideology1 <- lm(dom_percentage ~ ideology +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_is.republican1 <- lm(dom_percentage ~ is.republican +
                     as.factor(congress), 
               data = adver_cg)

adver_cg_chamber1 <- lm(dom_percentage ~ chamber +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_gender1 <- lm(dom_percentage ~ gender +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_seniority1 <- lm(dom_percentage ~ seniority +
                     as.factor(congress), 
                   data = adver_cg)


# legis
legis_cg_age1 <- lm(dom_percentage ~ age +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_ideology1 <- lm(dom_percentage ~ ideology +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_is.republican1 <- lm(dom_percentage ~ is.republican +
                     as.factor(congress), 
               data = legis_cg)

legis_cg_chamber1 <- lm(dom_percentage ~ chamber +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_gender1 <- lm(dom_percentage ~ gender +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_seniority1 <- lm(dom_percentage ~ seniority +
                     as.factor(congress), 
                   data = legis_cg)



# 2.2. output

# adver
modelsummary(
  list(
    "1" = adver_cg_age1,
    "2" = adver_cg_ideology1,
    "3" = adver_cg_is.republican1,
    "4" = adver_cg_gender1,
    "5" = adver_cg_chamber1,
    "6" = adver_cg_seniority1
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodel_adver_seperated_dom.html",
  escape = FALSE
)

# legis
modelsummary(
  list(
    "1" = legis_cg_age1,
    "2" = legis_cg_ideology1,
    "3" = legis_cg_is.republican1,
    "4" = legis_cg_gender1,
    "5" = legis_cg_chamber1,
    "6" = legis_cg_seniority1
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodel_legis_seperated_dom.html",
  escape = FALSE
)


# 2.3. robust check


# 批量生成线性模型的诊断图
plot_lm_diagnostics <- function(model_list, output_dir = "lm_diagnostics_plots") {
  
  if (!dir.exists(output_dir)) {
    dir.create(output_dir)
  }

  for (model_name in names(model_list)) {
    model <- model_list[[model_name]]
    
    # 1. QQ Plot
    p1 <- ggplot(model, aes(sample = .stdresid)) +
      stat_qq() + stat_qq_line(color = "red") +
      labs(title = paste0("QQ Plot - ", model_name),
           y = "Standardized Residuals") +
      theme_minimal()

    # 2. 残差 vs 拟合值
    df <- augment(model)
    p2 <- ggplot(df, aes(.fitted, .resid)) +
      geom_point(alpha = 0.6) +
      geom_smooth(method = "loess", se = FALSE, color = "blue") +
      geom_hline(yintercept = 0, linetype = "dashed", color = "red") +
      labs(title = paste0("Residuals vs Fitted - ", model_name),
           x = "Fitted values", y = "Residuals") +
      theme_minimal()

    # 3. Cook's Distance
    p3 <- ggplot(df, aes(seq_along(.cooksd), .cooksd)) +
      geom_bar(stat = "identity", fill = "tomato", alpha = 0.7) +
      labs(title = paste0("Cook's Distance - ", model_name),
           x = "Observation", y = "Cook's D") +
      theme_minimal()

    # 4. 杠杆图
    p4 <- ggplot(df, aes(.hat, .std.resid)) +
      geom_point(alpha = 0.6) +
      geom_smooth(method = "loess", se = FALSE, color = "blue") +
      labs(title = paste0("Leverage vs Std Residuals - ", model_name),
           x = "Leverage", y = "Standardized Residuals") +
      theme_minimal()

    # 合并图像
    combined_plot <- (p1 | p2) / (p3 | p4)

    # 保存
    ggsave(filename = file.path(output_dir, paste0("diagnostic_", model_name, ".png")),
           plot = combined_plot,
           width = 12, height = 10, dpi = 300)
  }

  cat("所有诊断图已保存至：", output_dir, "\n")
}

# 调用

# 构建模型列表

model_adver_lm_dom_multiple <- list(
  "adver_dom0" = adver_lm0,
  "adver_dom1" = adver_lm1,
  "adver_dom2" = adver_lm2,
  "adver_dom3" = adver_lm3,
  "adver_dom4" = adver_lm4
)

models_adver_lm_dom <- list(
  "age_dom" = adver_cg_age1,
  "ideology_dom" = adver_cg_ideology1,
  "is.republican_dom" = adver_cg_is.republican1,
  "gender_dom" = adver_cg_gender1,
  "chamber_dom" = adver_cg_chamber1,
  "seniority_dom" = adver_cg_seniority1
)

model_legis_lm_dom_multiple <- list(
  "legis_dom" = legis_lm0,
  "legis_dom1" = legis_lm1,
  "legis_dom2" = legis_lm2,
  "legis_dom3" = legis_lm3,
  "legis_dom4" = legis_lm4
)

models_legis_lm_dom <- list(
  "legis_age_dom" = legis_cg_age1,
  "legis_ideology_dom" = legis_cg_ideology1,
  "legis_is.republican_dom" = legis_cg_is.republican1,
  "legis_gender_dom" = legis_cg_gender1,
  "legis_chamber_dom" = legis_cg_chamber1,
  "legis_seniority_dom" = legis_cg_seniority1
)


# 运行函数
plot_lm_diagnostics(model_adver_lm_dom_multiple)
plot_lm_diagnostics(model_legis_lm_dom_multiple)
plot_lm_diagnostics(models_adver_lm_dom)
plot_lm_diagnostics(models_legis_lm_dom)

```


## 7.5 is. prof


```{r is.prof logit model}

## -----------------------------
# a. logit model - multiple variables
## -----------------------------

# 1. ridge & lasso ----

# adver
cv_lasso_adver <- run_glmnet_selection(adver_cg, is.prof ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 1,
  family = "binomial")
cv_ridge_adver <- run_glmnet_selection(adver_cg, is.prof ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0,
  family = "binomial")
cv_elnet_adver <- run_glmnet_selection(adver_cg, is.prof ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0.5,
  family = "binomial")

# legis
cv_lasso_legis <- run_glmnet_selection(legis_cg, is.prof ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 1,
  family = "binomial")
cv_ridge_legis <- run_glmnet_selection(legis_cg, is.prof ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0,
  family = "binomial")
cv_elnet_legis <- run_glmnet_selection(legis_cg, is.prof ~ age + party + ideology + chamber + gender + seniority + as.factor(congress), alpha = 0.5,
  family = "binomial")



# 2. model training ----

#adver
adver_logit0 <- glm(is.prof ~ age + is.republican + ideology + chamber + seniority + gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())


adver_logit_result <- analyze_logit_model(adver_logit0)


#legis
legis_logit0 <- glm(is.prof ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_result <- analyze_logit_model(legis_logit0)


# ajusted

adver_logit1 <- glm(is.prof ~ age +  chamber + ideology +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

legis_logit1 <- glm(is.prof ~ age +  chamber +ideology +  gender+as.factor(congress),  
                   data = legis_cg, 
                   family = binomial())

adver_logit2 <- glm(is.prof ~ age +  chamber + is.republican +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

legis_logit2 <- glm(is.prof ~ age +  chamber + is.republican +  gender+as.factor(congress),  
                   data = legis_cg, 
                   family = binomial())


adver_logit3 <- glm(is.prof ~ seniority +  chamber + is.republican +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

legis_logit3 <- glm(is.prof ~ seniority +  chamber + is.republican +  gender+as.factor(congress),  
                   data = legis_cg, 
                   family = binomial())


adver_logit4 <- glm(is.prof ~ seniority +  chamber + ideology +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

legis_logit4 <- glm(is.prof ~ seniority +  chamber + ideology +  gender+as.factor(congress),  
                   data = legis_cg, 
                   family = binomial())



# 4. check 

# pesudo r2
pR2(adver_logit1)
pR2(legis_logit1)

vif(adver_logit1)
vif(legis_logit1)


# robust check on legis_logit


# 5. output

modelsummary(
  list(
    "Adver All" = adver_logit0,
    "1" = adver_logit1,
    "2" = adver_logit2,
    "3" = adver_logit3,
    "4" = adver_logit4
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_adver_isprof.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)

modelsummary(
  list(
    "Legis All" = legis_logit0,
    "1" = legis_logit1,
    "2" = legis_logit2,
    "3" = legis_logit3,
    "4" = legis_logit4
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_legis_isprof.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)


## -----------------------------
# b. logit model - seperated variables
## -----------------------------

# 1.1 logit model construction

#adver
adver_logit_age <- glm(is.prof ~ age + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_ideology <- glm(is.prof ~ ideology + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_is.republican <- glm(is.prof ~ is.republican + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_gender <- glm(is.prof ~ gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_chamber <- glm(is.prof ~ chamber + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_logit_seniority <- glm(is.prof ~ seniority + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())


#legis
legis_logit_age <- glm(is.prof ~ age + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_ideology <- glm(is.prof ~ ideology + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_is.republican <- glm(is.prof ~ is.republican + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_gender <- glm(is.prof ~ gender + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_chamber <- glm(is.prof ~ chamber + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_logit_seniority <- glm(is.prof ~ seniority + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())


# 1.2. output

# adver
modelsummary(
  list(
    "1" = adver_logit_age,
    "2" = adver_logit_ideology,
    "3" = adver_logit_is.republican,
    "4" = adver_logit_gender,
    "5" = adver_logit_chamber,
    "6" = adver_logit_seniority
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_adver_seperated_isprof.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)


#legis
modelsummary(
  list(
    "1" = legis_logit_age,
    "2" = legis_logit_ideology,
    "3" = legis_logit_is.republican,
    "4" = legis_logit_gender,
    "5" = legis_logit_chamber,
    "6" = legis_logit_seniority
  ),
  fmt = 3,
  stars = TRUE,
  output = "logitmodel_legis_seperated_isprof.html",
  escape = FALSE,
  gof_omit = "AIC|BIC"
)



# 1.3. check 

# pesudo r2
pR2(adver_logit_age)
pR2(adver_logit_ideology)
pR2(adver_logit_is.republican)
pR2(adver_logit_gender)
pR2(adver_logit_seniority)
pR2(adver_logit_chamber)

pR2(legis_logit_age)
pR2(legis_logit_ideology)
pR2(legis_logit_is.republican)
pR2(legis_logit_gender)
pR2(legis_logit_seniority)
pR2(legis_logit_chamber)


vif(adver_logit_age)
vif(adver_logit_ideology)
vif(adver_logit_is.republican)
vif(adver_logit_gender)
vif(adver_logit_seniority)
vif(adver_logit_chamber)

vif(legis_logit_age)
vif(legis_logit_ideology)
vif(legis_logit_is.republican)
vif(legis_logit_gender)
vif(legis_logit_seniority)
vif(legis_logit_chamber)

# residual

model_list_adver_prof_logit <- list(
  "adver_prof_logit1" = adver_logit1,
  "adver_prof_logit2" = adver_logit2,
  "adver_prof_logit3" = adver_logit3,
  "adver_prof_logit4" = adver_logit4
)

model_list_legis_prof_logit <- list(
  "legis_prof_logit1" = legis_logit1,
  "legis_prof_logit2" = legis_logit2,
  "legis_prof_logit3" = legis_logit3,
  "legis_prof_logit4" = legis_logit4
)


plot_logit_residuals(model_list_adver_prof_logit, adver_cg)
plot_logit_residuals(model_list_legis_prof_logit, legis_cg)



```


## 7.6 prof_percentage

```{r prof_percentage lm model}


## -----------------------------
# a. regression model - mutilple variables
## -----------------------------

# 1. lasso/ ridge ----

# adver
cv_lasso_adver1 <- run_glmnet_selection(
  data = adver_cg,
  formula = prof_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 1,
  family = "gaussian"  
)

cv_ridge_adver1 <- run_glmnet_selection(
  data = adver_cg,
  formula = prof_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0,
  family = "gaussian"  
)

cv_elnet_adver1 <- run_glmnet_selection(
  data = adver_cg,
  formula = prof_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0.5,
  family = "gaussian"  
)

# legis
cv_lasso_legis1 <- run_glmnet_selection(
  data = legis_cg,
  formula = prof_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 1,
  family = "gaussian"  
)

cv_ridge_legis1 <- run_glmnet_selection(
  data = legis_cg,
  formula = prof_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0,
  family = "gaussian"  
)

cv_elnet_legis1 <- run_glmnet_selection(
  data = legis_cg,
  formula = prof_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress),
  alpha = 0.5,
  family = "gaussian"  
)


# 2. model training ----
# original
adver_lm0 <- lm(prof_percentage ~ age + 
                     is.republican + ideology + seniority+ 
                     chamber + gender + 
                     as.factor(congress), 
                   data = adver_cg)

legis_lm0 <- lm(prof_percentage ~ age + 
                     is.republican + ideology + seniority+ 
                     chamber + gender + 
                     as.factor(congress), 
                   data = legis_cg)

summary(adver_lm0)
summary(legis_lm0)

# adjusted

adver_lm1 <- lm(prof_percentage ~ age + 
                     is.republican + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


vif(lm(prof_percentage ~ age + is.republican + ideology + chamber + gender + seniority + as.factor(congress), data = adver_cg))
vif(lm(prof_percentage ~ age + is.republican + ideology + chamber + gender + as.factor(congress), data = adver_cg))
vif(lm(prof_percentage ~  seniority + is.republican + ideology + chamber + gender + as.factor(congress), data = adver_cg))



legis_lm1 <- lm(prof_percentage ~ age + 
                     chamber + gender + is.republican 
                     + as.factor(congress),   data = legis_cg )


adver_lm2 <- lm(prof_percentage ~ age + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


legis_lm2 <- lm(prof_percentage ~ age + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = legis_cg)


adver_lm3 <- lm(prof_percentage ~ seniority + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


legis_lm3 <- lm(prof_percentage ~ seniority + 
                     ideology + 
                     chamber + gender +
                     as.factor(congress), 
                   data = legis_cg)


adver_lm4 <- lm(prof_percentage ~ seniority + 
                     is.republican + 
                     chamber + gender +
                     as.factor(congress), 
                   data = adver_cg)


legis_lm4 <- lm(prof_percentage ~ seniority + 
                     is.republican + 
                     chamber + gender +
                     as.factor(congress), 
                   data = legis_cg)

# 3. qqplot check

# check basic assumptions of linear regression
#qqnorm(residuals(adver_cg_dom))
#qqline(residuals(adver_cg_dom), col = "red")
#mean(residuals(adver_cg_dom))
#vif(adver_cg_dom)

#qqnorm(residuals(legis_cg_dom))
#qqline(residuals(legis_cg_dom), col = "red")
#mean(residuals(legis_cg_dom))
#vif(legis_cg_dom)


# 4. output

modelsummary(
  list(
    "Adver LM" = adver_lm0,
    "1" = adver_lm1,
    "2" = adver_lm2,
    "3" = adver_lm3,
    "4" = adver_lm4
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodels_adver_prof.html",
  escape = FALSE
)

modelsummary(
  list(
    "Legis LM" = legis_lm0,
    "1" = legis_lm1,
    "2" = legis_lm2,
    "3" = legis_lm3,
    "4" = legis_lm4
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodels_legis_prof.html",
  escape = FALSE
)


## -----------------------------
# a. regression model - seperated variables
## -----------------------------


# 2.1. lm model construction

# adver
adver_cg_age1 <- lm(prof_percentage ~ age +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_ideology1 <- lm(prof_percentage ~ ideology +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_is.republican1 <- lm(prof_percentage ~ is.republican +
                     as.factor(congress), 
               data = adver_cg)

adver_cg_chamber1 <- lm(prof_percentage ~ chamber +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_gender1 <- lm(prof_percentage ~ gender +
                     as.factor(congress), 
                   data = adver_cg)

adver_cg_seniority1 <- lm(prof_percentage ~ seniority +
                     as.factor(congress), 
                   data = adver_cg)


# legis
legis_cg_age1 <- lm(prof_percentage ~ age +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_ideology1 <- lm(prof_percentage ~ ideology +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_is.republican1 <- lm(prof_percentage ~ is.republican +
                     as.factor(congress), 
               data = legis_cg)

legis_cg_chamber1 <- lm(prof_percentage ~ chamber +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_gender1 <- lm(prof_percentage ~ gender +
                     as.factor(congress), 
                   data = legis_cg)

legis_cg_seniority1 <- lm(prof_percentage ~ seniority +
                     as.factor(congress), 
                   data = legis_cg)



# 2.2. output

# adver
modelsummary(
  list(
    "1" = adver_cg_age1,
    "2" = adver_cg_ideology1,
    "3" = adver_cg_is.republican1,
    "4" = adver_cg_gender1,
    "5" = adver_cg_chamber1,
    "6" = adver_cg_seniority1
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodel_adver_seperated_prof.html",
  escape = FALSE
)

# legis
modelsummary(
  list(
    "1" = legis_cg_age1,
    "2" = legis_cg_ideology1,
    "3" = legis_cg_is.republican1,
    "4" = legis_cg_gender1,
    "5" = legis_cg_chamber1,
    "6" = legis_cg_seniority1
  ),
  fmt = 3,
  stars = TRUE,
  output = "lmmodel_legis_seperated_prof.html",
  escape = FALSE
)


# 2.3. robust check

# 构建模型列表

model_adver_lm_prof_multiple <- list(
  "adver_prof0" = adver_lm0,
  "adver_prof1" = adver_lm1,
  "adver_prof2" = adver_lm2,
  "adver_prof3" = adver_lm3,
  "adver_prof4" = adver_lm4
)

models_adver_lm_prof <- list(
  "age_prof" = adver_cg_age1,
  "ideology_prof" = adver_cg_ideology1,
  "is.republican_prof" = adver_cg_is.republican1,
  "gender_prof" = adver_cg_gender1,
  "chamber_prof" = adver_cg_chamber1,
  "seniority_prof" = adver_cg_seniority1
)

model_legis_lm_dom_multiple <- list(
  "legis_prof0" = legis_lm0,
  "legis_prof1" = legis_lm1,
  "legis_prof2" = legis_lm2,
  "legis_prof3" = legis_lm3,
  "legis_prof4" = legis_lm4
)

models_legis_lm_prof <- list(
  "legis_age_prof" = legis_cg_age1,
  "legis_ideology_prof" = legis_cg_ideology1,
  "legis_is.republican_prof" = legis_cg_is.republican1,
  "legis_gender_prof" = legis_cg_gender1,
  "legis_chamber_prof" = legis_cg_chamber1,
  "legis_seniority_prof" = legis_cg_seniority1
)


# 运行函数
plot_lm_diagnostics(model_adver_lm_dom_multiple)
plot_lm_diagnostics(model_legis_lm_dom_multiple)
plot_lm_diagnostics(models_adver_lm_prof)
plot_lm_diagnostics(models_legis_lm_prof)

```

### 7.7 plus independent variables 

```{r district-features-added}

# added variables 

adver_cg_clean <- adver_cg_clean %>%
  mutate(is.dom = ifelse(decision == "dom", 1, 0))

legis_cg_clean <- legis_cg_clean %>%
  mutate(is.dom = ifelse(decision == "dom", 1, 0))


adver_cg_clean <- adver_cg_clean %>% 
  mutate(is.republican = ifelse(party == "Republican", 1, 0))

legis_cg_clean <- legis_cg_clean %>% 
  mutate(is.republican = ifelse(party == "Republican", 1, 0))



names(adver_cg_clean)
names(legis_cg_clean)

unique(adver_cg_clean$ethnicity3)
unique(legis_cg_clean$ethnicity3)
unique(adver_cg_clean$payroll1)
unique(adver_cg_clean$payroll2)
unique(adver_cg_clean$payroll3)
unique(adver_cg_clean$employ1)
unique(adver_cg_clean$employ2)
unique(adver_cg_clean$employ3)

# 1. dependent variable == is.dom

# adver
cv_lasso_adver_plus <- run_glmnet_selection(adver_cg_clean, is.dom ~ age + is.republican + ideology + chamber + gender + seniority + ethnicity1 + payroll1 + employ1 + ppvi + npvi + as.factor(congress), alpha = 1,
  family = "binomial")
cv_ridge_adver_plus <- run_glmnet_selection(adver_cg_clean, is.dom ~ age + is.republican + ideology + chamber + gender + seniority + ethnicity1 + payroll1 + employ1 + ppvi + npvi + as.factor(congress), alpha = 0,
  family = "binomial")
cv_elnet_adver_plus <- run_glmnet_selection(adver_cg_clean, is.dom ~ age + is.republican + ideology + chamber + gender + seniority + ethnicity1 + payroll1 + employ1 + ppvi + npvi + as.factor(congress), alpha = 0.5,
  family = "binomial")

# legis
cv_lasso_legis_plus <- run_glmnet_selection(legis_cg_clean, is.dom ~ age + is.republican + ideology + chamber + gender + seniority + ethnicity1 + payroll1 + employ1 + ppvi + npvi + as.factor(congress), alpha = 1,
  family = "binomial")
cv_ridge_legis_plus <- run_glmnet_selection(legis_cg_clean, is.dom ~ age + is.republican + ideology + chamber + gender + seniority + ethnicity1 + payroll1 + employ1 + ppvi + npvi + as.factor(congress), alpha = 0,
  family = "binomial")
cv_elnet_legis_plus <- run_glmnet_selection(legis_cg_clean, is.dom ~ age + is.republican + ideology + chamber + gender + seniority + ethnicity1 + payroll1 + employ1 + ppvi + npvi + as.factor(congress), alpha = 0.5,
  family = "binomial")



adver_logit_plus <- glm(is.dom ~ ideology + ethnicity1 + payroll1 + employ1 + as.factor(congress), 
                   data = adver_cg_clean, 
                   family = binomial())


legis_logit_plus <- glm(is.dom ~ seniority + chamber + payroll1 + ethnicity1 + employ1 + ppvi + npvi + gender+as.factor(congress),  
                   data = legis_cg_clean, 
                   family = binomial())

summary(adver_logit_plus)
summary(legis_logit_plus)


modelsummary(
  list(
    "1" = adver_logit_plus,
    "2" = legis_logit_plus
  ),
  fmt = 3,
  stars = TRUE,
  output = "modelsummary_output_plus_logit.html",
  escape = FALSE
)


# 2. dependent variable == dom_percentage

# adver
adver_cg_payroll1 <- lm(dom_percentage ~ payroll1 +
                     as.factor(congress), 
                   data = adver_cg_clean)


adver_cg_employ1 <- lm(dom_percentage ~ employ1 +
                     as.factor(congress), 
                   data = adver_cg_clean)


adver_cg_ethnicity1 <- lm(dom_percentage ~ ethnicity1 +
                     as.factor(congress), 
                   data = adver_cg_clean)


adver_cg_ppvi <- lm(dom_percentage ~ ppvi +
                     as.factor(congress), 
                   data = adver_cg_clean)


adver_cg_npvi <- lm(dom_percentage ~ npvi +
                     as.factor(congress), 
                   data = adver_cg_clean)

modelsummary(
  list(
    "1" = adver_cg_payroll1,
    "2" = adver_cg_employ1,
    "3" = adver_cg_ethnicity1,
    "4" = adver_cg_ppvi,
    "5" = adver_cg_npvi
  ),
  fmt = 3,
  stars = TRUE,
  output = "modelsummary_output_plus_lm_adver.html",
  escape = FALSE
)


# legis
legis_cg_payroll1 <- lm(dom_percentage ~ payroll1 +
                     as.factor(congress), 
                   data = legis_cg_clean)


legis_cg_employ1 <- lm(dom_percentage ~ employ1 +
                     as.factor(congress), 
                   data = legis_cg_clean)


legis_cg_ethnicity1 <- lm(dom_percentage ~ ethnicity1 +
                     as.factor(congress), 
                   data = legis_cg_clean)


legis_cg_ppvi <- lm(dom_percentage ~ ppvi +
                     as.factor(congress), 
                   data = legis_cg_clean)


legis_cg_npvi <- lm(dom_percentage ~ npvi +
                     as.factor(congress), 
                   data = legis_cg_clean)

modelsummary(
  list(
    "1" = legis_cg_payroll1,
    "2" = legis_cg_employ1,
    "3" = legis_cg_ethnicity1,
    "4" = legis_cg_ppvi,
    "5" = legis_cg_npvi
  ),
  fmt = 3,
  stars = TRUE,
  output = "modelsummary_output_plus_lm_legis.html",
  escape = FALSE
)
```

### 7.8 plus entropy

```{r entropy-regression}

entropy_bills$congress <- as.character(entropy_bills$congress)
entropy_news$congress <- as.character(entropy_news$congress)

entropy_news_sel <- entropy_news %>% select(mid, congress, shannon_entropy)
entropy_bills_sel <- entropy_bills %>% select(mid, congress, shannon_entropy)

### Step 1: 合并 entropy_news 到 adver_cg
adver_cg_entropy <- adver_cg %>%
  left_join(entropy_news_sel, by = c("bioguide_id" = "mid", "congress" = "congress")) 

legis_cg_entropy <- legis_cg %>%
  left_join(entropy_bills_sel, by = c("bioguide_id" = "mid", "congress" = "congress"))


### Step 2: 回归模型设定
## (A) Logit: is.dom ~ entropy
adver_logit_entropy <- glm(is.dom ~ shannon_entropy + age + ideology + chamber + gender + as.factor(congress),
                   data = adver_cg_entropy,
                   family = binomial())

analyze_logit_model(adver_logit_entropy)

legis_logit_entropy <- glm(is.dom ~ shannon_entropy + age + ideology + chamber + gender + as.factor(congress),
                   data = legis_cg_entropy,
                   family = binomial())

analyze_logit_model(legis_logit_entropy)

## (B) OLS: dom_percentage ~ entropy
adver_ols_entropy <- lm(dom_percentage ~ shannon_entropy + age + ideology + chamber + gender + as.factor(congress),
                data = adver_cg_entropy)
summary(adver_ols_entropy)

legis_ols_entropy <- lm(dom_percentage ~ shannon_entropy + age + ideology + chamber + gender + as.factor(congress),
                data = legis_cg_entropy)
summary(legis_ols_entropy)


## (C) interact

adver_logit_entropy <- glm(is.dom ~  shannon_entropy*seniority + ideology + chamber + gender + as.factor(congress),
                   data = adver_cg_entropy,
                   family = binomial())

analyze_logit_model(adver_logit_entropy)

legis_logit_entropy <- glm(is.dom ~ shannon_entropy*seniority + ideology + chamber + gender + as.factor(congress),
                   data = legis_cg_entropy,
                   family = binomial())

analyze_logit_model(legis_logit_entropy)
interact_plot(legis_logit_entropy, pred = shannon_entropy, modx = seniority, interval = TRUE)

## (D) 分党派回归
### 民主党
model_dem <- lm(dom_percentage ~ shannon_entropy + age + ideology + chamber + gender + as.factor(congress),
                data = adver_cg_entropy %>% filter(party == "Democrat"))

### 共和党
model_rep <- lm(dom_percentage ~ shannon_entropy + age + ideology + chamber + gender + as.factor(congress),
                data = adver_cg_entropy %>% filter(party == "Republican"))

```


```{r interact-with-age}

adver_cg_entropy %>%
  select(age, seniority, shannon_entropy) %>%
  cor(use = "complete.obs")
cor.test(adver_cg_entropy$age, adver_cg_entropy$shannon_entropy)
cor.test(adver_cg_entropy$seniority, adver_cg_entropy$shannon_entropy)

legis_cg_entropy %>%
  select(age, seniority, shannon_entropy) %>%
  cor(use = "complete.obs")
cor.test(legis_cg_entropy$age, legis_cg_entropy$shannon_entropy)
cor.test(legis_cg_entropy$seniority, legis_cg_entropy$shannon_entropy)


```

```{r interact-with-gender}

cor(adver_cg_entropy$gender_numeric, adver_cg_entropy$shannon_entropy, method = "pearson")

adver_gender_entropy <- lm(shannon_entropy ~ age + ideology + chamber + gender + as.factor(congress),
                data = adver_cg_entropy)
summary(adver_gender_entropy)


legis_gender_entropy <- lm(shannon_entropy ~ age + ideology + chamber + gender + as.factor(congress),
                data = legis_cg_entropy)
summary(legis_gender_entropy)

```

# 8. improve model R2

## 8.1 change y

```{r use orginal decision as y}

library(nnet)

#adver
adver_decision <- multinom(decision ~ age +  ideology + chamber +  gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())

adver_decision_result <- analyze_logit_model(adver_decision)


#legis
legis_decision <- multinom(decision ~ age +  ideology + chamber +  gender + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())

legis_decision_result <- analyze_logit_model(legis_decision)

```


## 8.2 for logit model

### 8.2.1 interaction model


```{r interaction}

# 1. is. dom

# ideology * chamber and gender

adver_inter1 <- glm(is.dom ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = adver_cg, 
                   family = binomial())
analyze_logit_model(adver_inter1)



legis_inter1 <- glm(is.dom ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = legis_cg, 
                   family = binomial())
analyze_logit_model(legis_inter1)


# effects
install.packages("interactions")
library(interactions)

interact_plot(adver_inter1, pred = ideology, modx = chamber, interval = TRUE)
sim_slopes(adver_inter1, pred = ideology, modx = chamber)

interact_plot(adver_inter1, pred = ideology, modx = gender, interval = TRUE)
sim_slopes(adver_inter1, pred = ideology, modx = gender)

interact_plot(legis_inter1, pred = ideology, modx = chamber,interval = TRUE)
sim_slopes(legis_inter1, pred = ideology, modx = chamber)

interact_plot(legis_inter1, pred = ideology, modx = gender, interval = TRUE)
sim_slopes(legis_inter1, pred = ideology, modx = gender)


# combine


p1 <- interact_plot(adver_inter1, pred = ideology, modx = chamber, interval = TRUE) +
  ggtitle("Adver: Ideology × Chamber")

p2 <- interact_plot(adver_inter1, pred = ideology, modx = gender, interval = TRUE) +
  ggtitle("Adver: Ideology × Gender")

p3 <- interact_plot(legis_inter1, pred = ideology, modx = chamber, interval = TRUE) +
  ggtitle("Legis: Ideology × Chamber")

p4 <- interact_plot(legis_inter1, pred = ideology, modx = gender, interval = TRUE) +
  ggtitle("Legis: Ideology × Gender")


combined_plot <- (p1 | p2) / (p3 | p4)
combined_plot

ggsave("interaction_all.png", plot = combined_plot, width = 12, height = 8, dpi = 300)

```

### 8.2.2 spline

```{r spline}

# use gam()


adver_dom_gam <- gam(is.dom ~ s(age, k = 5) + s(ideology, k = 5) + chamber + gender +
                 as.factor(congress),
               family = binomial(),
               data = adver_cg)
summary(adver_dom_gam)

AIC(adver_dom_gam); MuMIn::r.squaredLR(adver_dom_gam) # pR2 works with glm only, but you can compute logLik etc
roc_gam <- roc(model.frame(adver_dom_gam)$is.dom, predict(adver_dom_gam, type = "response"))
auc(roc_gam)
plot(adver_dom_gam, pages = 1)  # 平滑函数可视化


# spline


adver_dom_spline <- glm(is.dom ~ ns(age, df=4) + ns(ideology, df=4) +
              chamber + gender + as.factor(congress),
    family = binomial(),
    data = adver_cg_clean)

AIC(adver_dom_spline); pR2(adver_dom_spline)["McFadden"]
MuMIn::r.squaredLR(adver_dom_spline)

legis_dom_spline <- glm(is.dom ~ ns(age, df = 4) + ns(ideology, df = 4) + chamber + gender + as.factor(congress),
                  family = binomial(), data = legis_cg)

AIC(legis_dom_spline); pR2(legis_dom_spline)["McFadden"]





```

### 8.2.3 poly

```{r poly-logit}

adver_cg$age2 <- adver_cg$age^2
adver_cg$ideo2 <- adver_cg$ideology^2

adver_cg_clean <- na.omit(adver_cg[, c("is.dom","age","age2","ideology","ideo2","chamber","gender","congress")])

adver_dom_poly <- glm(is.dom ~ age + age2 + ideology + ideo2 +
                      chamber + gender + as.factor(congress),
                      family = binomial(),
                      data = adver_cg_clean)

AIC(adver_dom_poly); pR2(adver_dom_poly)["McFadden"]


```


### 8.2.4 grouped

#### 8.2.4.1 grouped adver

```{r grouped-adver}

adver_cg_dem <- adver_cg %>% filter(party == "Democratic")
adver_cg_rep <- adver_cg %>% filter(party == "Republican")

# quartile

quantile(adver_cg_dem$ideology, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quantile(adver_cg_rep$ideology, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)

# ideology*chamber & ideology*gender

adver_inter_dem <- glm(is.dom ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = adver_cg_dem, 
                   family = binomial())
analyze_logit_model(adver_inter_dem)


adver_inter_rep <- glm(is.dom ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = adver_cg_rep, 
                   family = binomial())
analyze_logit_model(adver_inter_rep)

interact_plot(adver_inter_dem, pred = ideology, modx = chamber, data = adver_cg_dem)
interact_plot(adver_inter_rep, pred = ideology, modx = chamber, data = adver_cg_rep)

# combine -----

# 1. 创建预测数据框 
# ideology 从最小到最大
ideo_seq <- seq(min(c(adver_cg_dem$ideology, adver_cg_rep$ideology), na.rm=TRUE),
                max(c(adver_cg_dem$ideology, adver_cg_rep$ideology), na.rm=TRUE),
                length.out = 100)

# 民主党预测
pred_dem <- expand.grid(
  ideology = ideo_seq,
  chamber = c("Senate", "House of Representatives"),
  age = mean(adver_cg_dem$age, na.rm = TRUE),
  gender = "Female", # 基准
  congress = "115"
) %>%
  mutate(
    party = "Democrat",
    pred = predict(adver_inter_dem, newdata = ., type = "response")
  )

# 共和党预测
pred_rep <- expand.grid(
  ideology = ideo_seq,
  chamber = c("Senate", "House of Representatives"),
  age = mean(adver_cg_rep$age, na.rm = TRUE),
  gender = "Female", # 基准
  congress = "115"
) %>%
  mutate(
    party = "Republican",
    pred = predict(adver_inter_rep, newdata = ., type = "response")
  )

# 合并
pred_all <- bind_rows(pred_dem, pred_rep)


# 2. 画图 
p <- ggplot(pred_all, aes(x = ideology, y = pred, color = party, linetype = chamber)) +
  geom_line(size = 1) +
  scale_color_manual(values = c("Democrat" = "#0015BC", "Republican" = "#E9141D")) +
  scale_linetype_manual(values = c("Senate" = "solid", "House of Representatives" = "dashed")) +
  labs(
    x = "Ideology",
    y = "Predicted Probability of Domestic Mode",
    color = "Party",
    linetype = "Chamber",
    title = "Ideology × Chamber Interaction by Party"
  ) +
  theme_minimal(base_size = 14)

ggsave("interaction_plot.png", plot = p, width = 8, height = 6, dpi = 300)

# combine 2----

install.packages("marginaleffects")
library(marginaleffects)

# 民主党
pred_dem <- avg_predictions(
  adver_inter_dem,
  variables = list(
    ideology = seq(min(adver_cg_dem$ideology, na.rm=TRUE),
                   max(adver_cg_dem$ideology, na.rm=TRUE),
                   length.out=50),
    chamber = c("House of Representatives","Senate"),
    gender = c("Male","Female")
  ),
  by = c("ideology", "chamber", "gender")
) %>% 
  mutate(party = "Democrat")

# 共和党
pred_rep <- avg_predictions(
  adver_inter_rep,
  variables = list(
    ideology = seq(min(adver_cg_rep$ideology, na.rm=TRUE),
                   max(adver_cg_rep$ideology, na.rm=TRUE),
                   length.out=50),
    chamber = c("House of Representatives","Senate"),
    gender = c("Male","Female")
  ),
  by = c("ideology", "chamber", "gender")
) %>% 
  mutate(party = "Republican")

# 合并
pred_all <- bind_rows(pred_dem, pred_rep)

p_all <- ggplot(pred_all, aes(x = ideology, y = estimate,
                     color = interaction(gender, chamber),
                     linetype = chamber)) +
  geom_line(size=1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = interaction(gender, chamber)),
              alpha = 0.2, color = NA) +
  facet_wrap(~party) +
  labs(
    x = "Ideology (liberal → conservative)",
    y = "Predicted Probability of Domestic Mode",
    color = "Gender × Chamber",
    fill = "Gender × Chamber",
    linetype = "Chamber"
  ) +
  theme_minimal(base_size = 14)

ggsave("interaction_all_party&gender&chamber.png", plot = p_all, width = 12, height = 4, dpi = 300)

```


#### 8.2.4.2 grouped legis

```{r grouped-legis}

legis_cg_dem <- legis_cg %>% filter(party == "Democratic")
legis_cg_rep <- legis_cg %>% filter(party == "Republican")

quantile(adver_cg_dem$ideology, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)
quantile(adver_cg_rep$ideology, probs = c(0, 0.25, 0.5, 0.75, 1), na.rm = TRUE)

# export
wb <- wb_workbook()
wb$add_worksheet("dem"); wb$add_data("dem", legis_cg_dem)
wb$add_worksheet("rep"); wb$add_data("rep", legis_cg_rep)
wb$add_worksheet("dem_adver"); wb$add_data("dem_adver", adver_cg_dem)
wb$add_worksheet("rep_adver"); wb$add_data("rep_adver", adver_cg_rep)
wb_save(wb, "cg_grouped.xlsx")

# ideology*chamber & ideology*gender

legis_inter_dem <- glm(is.dom ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = legis_cg_dem, 
                   family = binomial())
analyze_logit_model(legis_inter_dem)


legis_inter_rep <- glm(is.dom ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = legis_cg_rep, 
                   family = binomial())
analyze_logit_model(legis_inter_rep)

interact_plot(legis_inter_dem, pred = ideology, modx = chamber, data = legis_cg_dem)
interact_plot(legis_inter_rep, pred = ideology, modx = chamber, data = legis_cg_rep)


# combine----

# 民主党
pred_dem <- avg_predictions(
  legis_inter_dem,
  variables = list(
    ideology = seq(min(legis_cg_dem$ideology, na.rm=TRUE),
                   max(legis_cg_dem$ideology, na.rm=TRUE),
                   length.out=50),
    chamber = c("House of Representatives","Senate"),
    gender = c("Male","Female")
  ),
  by = c("ideology", "chamber", "gender")
) %>% 
  mutate(party = "Democrat")

# 共和党
pred_rep <- avg_predictions(
  legis_inter_rep,
  variables = list(
    ideology = seq(min(legis_cg_rep$ideology, na.rm=TRUE),
                   max(legis_cg_rep$ideology, na.rm=TRUE),
                   length.out=50),
    chamber = c("House of Representatives","Senate"),
    gender = c("Male","Female")
  ),
  by = c("ideology", "chamber", "gender")
) %>% 
  mutate(party = "Republican")

# 合并
pred_all <- bind_rows(pred_dem, pred_rep)

p_all <- ggplot(pred_all, aes(x = ideology, y = estimate,
                     color = interaction(gender, chamber),
                     linetype = chamber)) +
  geom_line(size=1) +
  geom_ribbon(aes(ymin = conf.low, ymax = conf.high, fill = interaction(gender, chamber)),
              alpha = 0.2, color = NA) +
  facet_wrap(~party) +
  labs(
    x = "Ideology (liberal → conservative)",
    y = "Predicted Probability of Domestic Mode",
    color = "Gender × Chamber",
    fill = "Gender × Chamber",
    linetype = "Chamber"
  ) +
  theme_minimal(base_size = 14)

ggsave("interaction_all_party&gender&chamber_legis.png", plot = p_all, width = 12, height = 4, dpi = 300)




```




## 8.3 lm model(higher r2)

### 8.3.1 interaction model

#### 8.3.1.1 adver

```{r interaction-adver}

# 1. dom_percentage

# ideology * chamber and gender

adver_lm_inter1 <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress),
                   data = adver_cg)
summary(adver_lm_inter1)


legis_lm_inter1 <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress),
                   data = legis_cg)
summary(legis_lm_inter1)

p1 <- interact_plot(adver_lm_inter1, pred = ideology, modx = gender, data = adver_cg)
ggsave("gendered_interaction_plot.png", plot = p1, width = 8, height = 6, dpi = 300)



# effects ----

interact_plot(adver_lm_inter1, pred = ideology, modx = chamber,
              plot.points = TRUE, interval = TRUE)
interact_plot(adver_lm_inter1, pred = ideology, modx = gender,
              plot.points = TRUE, interval = TRUE)


interact_plot(legis_lm_inter1, pred = ideology, modx = chamber,
              plot.points = TRUE, interval = TRUE)
interact_plot(legis_lm_inter1, pred = ideology, modx = gender,
              plot.points = TRUE, interval = TRUE)


# grouped effects ----

adver_inter_dem_lm <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = adver_cg_dem)
summary(adver_inter_dem_lm)

adver_inter_rep_lm <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = adver_cg_rep)
summary(adver_inter_rep_lm)


interact_plot(adver_inter_dem_lm, pred = ideology, modx = chamber, data = adver_cg_dem)
sim_slopes(adver_inter_dem_lm, pred = ideology, modx = chamber)

interact_plot(adver_inter_rep_lm, pred = ideology, modx = chamber, data = adver_cg_rep)
sim_slopes(adver_inter_rep_lm, pred = ideology, modx = chamber)

interact_plot(adver_inter_rep_lm, pred = ideology, modx = gender, data = adver_cg_rep)
sim_slopes(adver_inter_rep_lm, pred = ideology, modx = gender)


# output result together with is.dom

modelsummary(
  list(
    "is.dom(full)" = adver_inter1,
    "is.dom(dem)" = adver_inter_dem,
    "is.dom(rep)" = adver_inter_rep,
    "dom_percentage(full)" = adver_lm_inter1,
    "dom_percentage(dem)" = adver_inter_dem_lm,
    "dom_percentage(rep)" = adver_inter_rep_lm
  ),
  fmt = 3,
  stars = TRUE,
  output = "adver_interaction.html",
  escape = FALSE
)
```


#### 8.3.1.2 legis

```{r interaction-legis}

# 1. dom_percentage

# ideology * chamber and gender

legis_lm_inter1 <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress),
                   data = legis_cg)
summary(legis_lm_inter1)


legis_lm_inter1 <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress),
                   data = legis_cg)
summary(legis_lm_inter1)

p1 <- interact_plot(legis_lm_inter1, pred = ideology, modx = gender, data = legis_cg)
ggsave("gendered_interaction_plot.png", plot = p1, width = 8, height = 6, dpi = 300)

sim_slopes(legis_lm_inter1, pred = ideology, modx = gender)

# effects ----

interact_plot(legis_lm_inter1, pred = ideology, modx = chamber, interval = TRUE)
interact_plot(legis_lm_inter1, pred = ideology, modx = gender,interval = TRUE)


interact_plot(legis_lm_inter1, pred = ideology, modx = chamber, interval = TRUE)
interact_plot(legis_lm_inter1, pred = ideology, modx = gender,interval = TRUE)


# grouped effects ----

legis_inter_dem_lm <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = legis_cg_dem)
summary(legis_inter_dem_lm)

legis_inter_rep_lm <- lm(dom_percentage ~ age + ideology*chamber + ideology*gender + as.factor(congress), 
                   data = legis_cg_rep)
summary(legis_inter_rep_lm)


interact_plot(legis_inter_dem_lm, pred = ideology, modx = chamber, data = legis_cg_dem)
sim_slopes(legis_inter_dem_lm, pred = ideology, modx = chamber)

interact_plot(legis_inter_rep_lm, pred = ideology, modx = chamber, data = legis_cg_rep)
sim_slopes(legis_inter_rep_lm, pred = ideology, modx = chamber)

interact_plot(legis_inter_rep_lm, pred = ideology, modx = gender, data = legis_cg_rep)
sim_slopes(legis_inter_rep_lm, pred = ideology, modx = gender)




# output result together with is.dom

modelsummary(
  list(
    "is.dom(full)" = legis_inter1,
    "is.dom(dem)" = legis_inter_dem,
    "is.dom(rep)" = legis_inter_rep,
    "dom_percentage(full)" = legis_lm_inter1,
    "dom_percentage(dem)" = legis_inter_dem_lm,
    "dom_percentage(rep)" = legis_inter_rep_lm
  ),
  fmt = 3,
  stars = TRUE,
  output = "legis_interaction.html",
  escape = FALSE
)


```




